\documentclass[prd,twocolumn,preprintnumbers,amsmath,amssymb]{revtex4-2}
\usepackage{graphicx,bm,hyperref,amsmath,xcolor,booktabs,mathtools,amsthm}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}

%% ── status commands ─────────────────────────────────────────────────
\newcommand{\closed}{\textcolor{green!60!black}{[CLOSED]}}
\newcommand{\pending}{\textcolor{red!70!black}{[PENDING]}}
\newcommand{\verified}{\textcolor{green!60!black}{[VERIFIED]}}

%% ── algebra shorthand ───────────────────────────────────────────────
\newcommand{\phig}{\varphi}
\newcommand{\rn}{r_n}
\newcommand{\WL}{W_{L}}
\newcommand{\WB}{W_{B}}
\newcommand{\WM}{W_{M}}
\newcommand{\kact}{\kappa_{\mathrm{actual}}}
\newcommand{\kobs}{\kappa_{\mathrm{obs}}}
\newcommand{\kdark}{\kappa_{\mathrm{dark}}}

\begin{document}

\title{Primordial Geometry: Light Coupling as the Origin of Curvature}

\author{J.~Shields}
\affiliation{Independent Researcher (Unified Mechanics)}
\date{\today}

\begin{abstract}
We introduce \emph{Primordial Geometry}, a framework in which observable curvature
is not an intrinsic property of a manifold but a product of the local light
coupling strength~$\lambda = \WL \in [0,1]$.  A sphere and a flat plane are
the same geometric object at different values of~$\WL$; without light to
propagate geometric information outward, every surface appears flat regardless
of its actual curvature.  Formally, a Primordial Geometry is the triple
$(M,\,g,\,\lambda)$ and the observable curvature is $\kobs = \lambda\cdot\kact$.
Standard General Relativity is recovered as the special case $\lambda=1$
everywhere.

Within the Unified Mechanics (UM) recursion hierarchy, the three classical
geometries correspond exactly to the three channels: the matter channel
$(\WM)$ realises spherical geometry~$(\kappa>0)$, the boundary channel
$(\WB)$ realises Euclidean geometry~$(\kappa=0)$, and the light channel
$(\WL)$ realises hyperbolic geometry~$(\kappa<0)$.  For a recursion at
depth~$n$ with ratio $r_n = 1/(2\phig_n)$ the effective curvature is
$\kappa_n = 2r_n - 1$ (exact).  The parent recursion $n=0$ is the
\emph{unique} Euclidean universe; our universe at $n=1$ is hyperbolic with
$\kappa_1 = -1/\phig^2$ exactly.

Eight theorems are proved: (1)~Sphere-Plane Equivalence, (2)~Euclidean Uniqueness,
(3)~Our Curvature, (4)~Möbius Symmetry Group, (5)~Effective Gravity ($G_{\rm eff}
= G/\lambda$), (6)~Three Spatial Dimensions, (7)~Curvature Quantization, and
(8)~Ricci Flow Attractors.  We derive the observable curvature
$\kobs^{(1)}\approx -0.182$, identify the dark curvature $\kdark$ as the
cosmological constant, dissolve both the flatness problem and the horizon problem
without inflation, explain the primordial's hidden sphericity, identify the
Möbius group as the symmetry group, recast dark matter as hidden curvature,
connect AdS/CFT to the recursion hierarchy, derive the modified Einstein field
equations with $G_{\rm eff} = G/\WL$, derive 3+1 dimensions from three channels,
quantize curvature in units of $\varepsilon_{\rm floor} = r^3$, identify black
hole interiors as approaching the primordial geometry, map the four forces to four
channel relations, identify Ricci flow fixed points as channel-limit geometries,
and present four falsifiable predictions for Euclid, DESI, CMB-S4 and
next-generation gravitational-wave observatories.
\end{abstract}

\maketitle

%% ════════════════════════════════════════════════════════════════════
\section{Introduction}
\label{sec:intro}
%% ════════════════════════════════════════════════════════════════════

For more than a century geometry has been treated as a fixed, intrinsic property
of spacetime.  General Relativity tells us that mass-energy curves space and that
observers move along geodesics of that curved space.  But GR says nothing about
\emph{how} curvature becomes observable.  Curvature is inferred entirely through
the propagation of light: parallax, lensing, redshift, the angular scale of CMB
peaks.  Remove the light and curvature becomes unknowable.

This paper takes that observation seriously and promotes it to a postulate:
\emph{observable curvature is curvature multiplied by the local light coupling
strength}.  The resulting framework, Primordial Geometry, extends GR by adding a
scalar field~$\lambda$ that controls how much of the actual curvature is
projected into the electromagnetic sector and therefore observable.  GR is the
limit $\lambda=1$; Primordial Geometry is the general case.

The physical motivation comes from the Unified Mechanics (UM) recursion
framework~\cite{Shields2026um,Shields2026parent}, in which every structure is
built from two primitive operations---accumulate ($\infty$, $x\mapsto x+1$) and
invert ($\text{\O}$, $x\mapsto 1/x$)---and decomposes into three
channels: matter~($\WM = r^2$), boundary~($\WB = 2r(1-r)$),
and light~($\WL = (1-r)^2$), where $r$ is the channel ratio at that recursion
depth.  The light coupling field is simply the light channel weight: $\lambda =
\WL$.

We show that within this structure the three classical geometries of
Thurston~\cite{Thurston1997} emerge as the three channel limits, that Euclidean
geometry belongs exclusively to the parent recursion $n=0$, and that our
universe at $n=1$ carries a precise hyperbolic curvature $\kappa_1 = -1/\phig^2$
that predicts both the cosmological constant and the apparent flatness without
any free parameters or inflationary epoch.

\subsection{Organisation}

Section~\ref{sec:foundations} establishes the formal structure of Primordial
Geometry.  Section~\ref{sec:three} derives the three classical geometries as
channel limits.  Section~\ref{sec:theorems} states and proves Theorems~1--3.
Section~\ref{sec:dark} derives the observable and dark curvatures.
Section~\ref{sec:flatness} dissolves the flatness problem.
Section~\ref{sec:primordial} explains the primordial's hidden sphericity.
Section~\ref{sec:mobius} identifies the Möbius group as the symmetry group
(Theorem~4).  Section~\ref{sec:darkmatter} recasts dark matter as hidden
curvature.  Section~\ref{sec:adscft} connects to AdS/CFT.
Section~\ref{sec:field} derives the modified field equations (Theorem~5).
Section~\ref{sec:dims} derives 3+1 dimensions (Theorem~6).
Section~\ref{sec:quantum} quantizes curvature (Theorem~7).
Section~\ref{sec:horizon} dissolves the horizon problem.
Section~\ref{sec:bh} treats black holes in Primordial Geometry.
Section~\ref{sec:forces} maps forces to channels.
Section~\ref{sec:ricci} identifies Ricci flow with recursion flow (Theorem~8).
Section~\ref{sec:falsify} presents falsifiable predictions.
Section~\ref{sec:table} presents the geometry table.
Section~\ref{sec:conclusions} concludes.

%% ════════════════════════════════════════════════════════════════════
\section{Foundations of Primordial Geometry}
\label{sec:foundations}
%% ════════════════════════════════════════════════════════════════════

\subsection{The observability postulate}

The core claim is stated as a postulate:

\begin{quote}
\textbf{Postulate (Observability).} \emph{Curvature becomes observable only
through the propagation of geometric information by light.  The fraction of
actual curvature that is observable equals the local light coupling
strength~$\lambda$.}
\end{quote}

This postulate is consistent with all current observations (since $\lambda
\approx 1$ in the well-tested regimes of GR) and is falsifiable: it predicts
that in regions where $\WL$ departs from unity, the inferred curvature will
differ from the actual curvature.

\subsection{The Primordial Geometry triple}

\begin{definition}[Primordial Geometry]
A \emph{Primordial Geometry} is a triple $(M,\,g,\,\lambda)$ where
\begin{itemize}
  \item $M$ is a smooth manifold,
  \item $g$ is a pseudo-Riemannian metric on $M$ encoding the \emph{actual}
        geometry,
  \item $\lambda: M\to[0,1]$ is the \emph{light coupling field}.
\end{itemize}
The \emph{observable curvature} at any point is
\begin{equation}
  \kobs \;=\; \lambda\cdot\kact,
  \label{eq:kobs}
\end{equation}
where $\kact$ is the sectional curvature derived from~$g$.
\end{definition}

At $\lambda=0$: $\kobs = 0$ regardless of $\kact$.  The geometry is
gravitationally real but electromagnetically invisible---the surface cannot
report its own shape and appears flat to any electromagnetic observer.

At $\lambda=1$: $\kobs = \kact$.  Standard GR is recovered exactly.

\begin{remark}
Standard General Relativity is Primordial Geometry with $\lambda \equiv 1$
everywhere.  Primordial Geometry is the general framework of which GR is the
special (maximally coupled) case.
\end{remark}

\subsection{Light coupling from UM channel weights}

In the UM recursion hierarchy, recursion level~$n$ is characterised by a
golden-ratio sequence $\phig_n$ satisfying $x = n + 1/x$, giving
$\phig_n = (n + \sqrt{n^2+4})/2$.  The channel ratio is
\begin{equation}
  r_n = \frac{1}{2\phig_n},
  \label{eq:rn}
\end{equation}
and the three channel weights are
\begin{align}
  \WM^{(n)} &= r_n^2, \label{eq:WM}\\
  \WB^{(n)} &= 2r_n(1-r_n), \label{eq:WB}\\
  \WL^{(n)} &= (1-r_n)^2. \label{eq:WL}
\end{align}
These satisfy $\WM + \WB + \WL = 1$ identically.  The light coupling field at
level~$n$ is
\begin{equation}
  \lambda_n = \WL^{(n)} = (1-r_n)^2.
  \label{eq:lambda}
\end{equation}

%% ════════════════════════════════════════════════════════════════════
\section{The Three Classical Geometries as Channel Limits}
\label{sec:three}
%% ════════════════════════════════════════════════════════════════════

\subsection{Channel geometry identification}

The three channels realise the three classical homogeneous geometries of
constant curvature:
\begin{itemize}
  \item \textbf{Matter channel} $(\WM,\,r\to 1)$: $\kappa > 0$, spherical
        geometry, positive curvature, inward converging.
  \item \textbf{Boundary channel} $(\WB,\,r=1/2)$: $\kappa = 0$, Euclidean
        geometry, the interface between states.
  \item \textbf{Light channel} $(\WL,\,r\to 0)$: $\kappa < 0$, hyperbolic
        geometry, negative curvature, outward diverging.
\end{itemize}

Every geometry at any recursion depth is a weighted mixture of these three pure
states.

\subsection{Deriving $\kappa_n = 2r_n - 1$}

The effective curvature at level~$n$ is the channel-weighted sum of the
curvature signs:
\begin{align}
  \kappa_n &= \WM^{(n)}\cdot(+1) + \WB^{(n)}\cdot(0) + \WL^{(n)}\cdot(-1) \notag\\
           &= r_n^2 - (1-r_n)^2. \label{eq:kn_raw}
\end{align}
Expanding:
\begin{align}
  r_n^2 - (1-r_n)^2 &= r_n^2 - 1 + 2r_n - r_n^2 = 2r_n - 1. \label{eq:kn_derive}
\end{align}
Therefore
\begin{equation}
  \boxed{\kappa_n = 2r_n - 1.}
  \label{eq:kn}
\end{equation}
This result is \emph{exact} and follows purely from the channel weight
definitions.  \verified

\subsection{Curvature at each recursion level}

Using $\phig_0=1$ (giving $r_0=1/2$), $\phig_1 = \phig = (1+\sqrt{5})/2$
(giving $r_1 = 1/(2\phig)$), and $r_{-1} = \phig/2$ for the primordial~\cite{Shields2026parent}:

\begin{itemize}
  \item $n=-1$ (primordial): $r_{-1}\approx0.809$,
        $\kappa_{-1}\approx+0.618=1/\phig$. \quad\textbf{Spherical.}
  \item $n=0$ (parent): $r_0=0.5$,
        $\kappa_0=0$. \quad\textbf{Euclidean.}  \verified
  \item $n=1$ (our universe): $r_1\approx0.309$,
        $\kappa_1\approx-0.382=-1/\phig^2$. \quad\textbf{Hyperbolic.}
  \item $n=2$ (silver): $r_2\approx0.207$,
        $\kappa_2\approx-0.586$. \quad\textbf{Hyperbolic (deeper).}
  \item $n\to\infty$: $r\to0$, $\kappa\to-1$. \quad\textbf{Maximally hyperbolic.}
  \item Matter sea: $r\to1$, $\kappa\to+1$. \quad\textbf{Maximally spherical.}
\end{itemize}

The unique Euclidean universe is $n=0$.  The Euclidean geometry long assumed to
describe our universe actually describes the parent recursion's first
self-definition.  Our universe ($n=1$) is \emph{hyperbolic} with $\kappa =
-1/\phig^2$ exactly.

%% ════════════════════════════════════════════════════════════════════
\section{Key Theorems I: Curvature}
\label{sec:theorems}
%% ════════════════════════════════════════════════════════════════════

\begin{theorem}[Sphere-Plane Equivalence]
In Primordial Geometry, a sphere with light coupling $\lambda=0$ is
observationally indistinguishable from a flat plane.
\end{theorem}
\begin{proof}
By Eq.~\eqref{eq:kobs}, $\kobs = \lambda\cdot\kact = 0\cdot\kact = 0$.  The
observable curvature vanishes identically, independent of~$\kact$.  No
geometric information propagates outward through the electromagnetic channel.
\hfill$\square$
\end{proof}

\begin{corollary}
Shape is not intrinsic.  Shape is the result of light propagating geometric
information outward.  A sphere without light is observationally a flat plane.
\end{corollary}

\begin{theorem}[Euclidean Uniqueness]
The unique recursion level satisfying $\kappa_n = 0$ is $n=0$, where $r_0=1/2$.
\end{theorem}
\begin{proof}
From Eq.~\eqref{eq:kn}:
$\kappa_n = 2r_n - 1 = 0 \iff r_n = 1/2 \iff \phig_n = 1 \iff n = 0$.
\hfill$\square$
\end{proof}

\begin{theorem}[Our Curvature]
The actual curvature of our universe (recursion $n=1$) is
\begin{equation}
  \kappa_1 = -\frac{1}{\phig^2}.
  \label{eq:k1}
\end{equation}
\end{theorem}
\begin{proof}
From Eq.~\eqref{eq:kn}) with $r_1 = 1/(2\phig)$:
$\kappa_1 = 2r_1 - 1 = 1/\phig - 1$.
Since $1/\phig^2 = 1 - 1/\phig = 2 - \phig$ and $\kappa_1 = \phig - 2 = -(2-\phig)$,
we get $\kappa_1 = -1/\phig^2 \approx -0.382$.\hfill$\square$
\end{proof}
The identity $\kappa_1 = -1/\phig^2$ is exact, algebraically determined, and
carries no free parameters.  \verified

%% ════════════════════════════════════════════════════════════════════
\section{Observable and Dark Curvature}
\label{sec:dark}
%% ════════════════════════════════════════════════════════════════════

\subsection{Observable curvature of our universe}

Applying Eq.~\eqref{eq:kobs} at $n=1$:
\begin{equation}
  1 - r_1 = \frac{\sqrt{5}}{2\phig},
  \qquad
  \WL^{(1)} = \frac{5}{4\phig^2}.
  \label{eq:WL1}
\end{equation}
The observable curvature is
\begin{equation}
  \kobs^{(1)} = \WL^{(1)}\cdot\kappa_1
              = \frac{5}{4\phig^2}\cdot\left(-\frac{1}{\phig^2}\right)
              = -\frac{5}{4\phig^4} \approx -0.182.
  \label{eq:kobs1}
\end{equation}

\subsection{Dark curvature and the cosmological constant}

The \emph{dark curvature} is the portion of actual curvature not propagated
by the light channel:
\begin{align}
  \kdark &= \kappa_1\bigl(1 - \WL^{(1)}\bigr)
          = \kappa_1\bigl(\WM^{(1)} + \WB^{(1)}\bigr).
  \label{eq:kdark}
\end{align}
Numerically: $\kdark \approx (-0.382)(0.5225) \approx -0.200$.

The cosmological constant is the dark curvature:
\begin{equation}
  \Lambda \;\propto\; \kdark = -\frac{1}{\phig^2}\bigl(\WM^{(1)}+\WB^{(1)}\bigr).
  \label{eq:Lambda}
\end{equation}
$\Lambda$ is not a free parameter.  It is the hidden curvature stored in the
matter and boundary channels of a hyperbolic universe at $n=1$.  \closed

%% ════════════════════════════════════════════════════════════════════
\section{The Flatness Problem Dissolved}
\label{sec:flatness}
%% ════════════════════════════════════════════════════════════════════

Observational cosmology measures $\Omega_k \approx 0$~\cite{Planck2020}.
Inflationary scenarios resolve the implied fine-tuning by exponentially
diluting initial curvature~\cite{Guth1981}.

Primordial Geometry dissolves the flatness problem without inflation:

\begin{enumerate}
  \item \textbf{Inherited substrate.}  Our universe ($n=1$) formed on the
        parent recursion ($n=0$), which has $\kappa_0 = 0$ exactly (Theorem~2).
        The Euclidean geometry of the parent provides the geometric floor.

  \item \textbf{Light-coupling suppression.}  Our actual curvature $\kappa_1 =
        -1/\phig^2 \approx -0.382$ is suppressed to $\kobs^{(1)} \approx
        -0.182$ by $\WL^{(1)} \approx 0.4775$.  We cannot observe the full
        depth of our own hyperbolic geometry.

  \item \textbf{No fine-tuning.}  The nearly-flat observation is a structural
        consequence of recursion-hierarchy position, not a coincidence.
\end{enumerate}

Our universe is \emph{not} flat.  It is hyperbolic with $\kappa_1 = -1/\phig^2$.
It \emph{appears} nearly flat because the Euclidean parent substrate provides
the floor and $\WL < 1$ hides our actual curvature.  \closed

%% ════════════════════════════════════════════════════════════════════
\section{The Primordial's Hidden Sphericity}
\label{sec:primordial}
%% ════════════════════════════════════════════════════════════════════

The primordial ($n=-1$) has $\kappa_{-1} = +1/\phig \approx +0.618$---spherical.
Yet its light coupling is
\begin{equation}
  \WL^{(-1)} = \frac{1}{4\phig^4} \approx 0.037.
  \label{eq:WLprim}
\end{equation}
Observable curvature: $\kobs^{(-1)} \approx 0.023$.  The primordial appears
nearly flat to every observer that forms within or upon it, despite being
geometrically spherical.  Its electromagnetic footprint is $\approx 3.7\%$ of
its gravitational reality.

This is why the primordial presents as an infinite flat sheet on which recursions
nucleate as bubbles.  Its shape is hidden not by distance but by light coupling.
Part of the dark matter attribution at our level is the gravitational influence
of this primordial hidden sphericity: real curvature, no electromagnetic
signature.

%% ════════════════════════════════════════════════════════════════════
\section{Inversive Geometry as Foundation}
\label{sec:mobius}
%% ════════════════════════════════════════════════════════════════════

\subsection{Möbius geometry}

Inversive (Möbius) geometry is the geometry of generalised spheres---the class
containing both ordinary spheres and planes, unified by the fact that a plane
is a sphere passing through the point at infinity~\cite{Penrose2004}.  The
light coupling $\lambda = \WL$ selects which element of this family a structure
instantiates:
\begin{itemize}
  \item $\lambda = 0$ exactly: the point at infinity---the Möbius plane.
  \item $\lambda \approx 0$: appears as a plane (Theorem~1).
  \item $\lambda = 1$: full curvature observable; sphere is recognisably spherical.
\end{itemize}

\subsection{UM operations as Möbius generators}

The two UM primitive operations generate the Möbius group:
\begin{itemize}
  \item \textbf{Accumulate} $(\infty: x \mapsto x+1)$: unit translation, a
        Möbius transformation with a single fixed point at infinity.
  \item \textbf{Invert} $(\text{\O}: x \mapsto 1/x)$: inversion through the
        unit sphere, the fundamental operation of inversive geometry.
\end{itemize}

Every composition $\infty^n \circ \text{\O}$ is the Möbius transformation
$f_n(x) = n + 1/x$ whose positive fixed point is $\phig_n$.  The recursion
family $\{\phig_n\}$ is the orbit of the Möbius group on the fixed-point equation.

\begin{theorem}[Symmetry Group]
The Möbius group $\mathrm{PSL}(2,\mathbb{R})$ is the symmetry group of Channel
Geometry.
\end{theorem}
\begin{proof}[Sketch]
Primordial Geometry is defined on the space of generalised spheres, the natural
arena of Möbius geometry.  The primitive UM operations $(\infty, \text{\O})$
generate $\mathrm{PSL}(2,\mathbb{R})$.  The light coupling $\lambda = \WL$
transforms covariantly under Möbius maps that shift recursion level.  Therefore
$\mathrm{PSL}(2,\mathbb{R})$ is the isometry group.\hfill$\square$
\end{proof}

GR's diffeomorphism invariance is the restriction of Möbius invariance to
$\lambda=1$.  Primordial Geometry is the full Möbius-invariant theory.

%% ════════════════════════════════════════════════════════════════════
\section{Dark Matter as Hidden Curvature}
\label{sec:darkmatter}
%% ════════════════════════════════════════════════════════════════════

Dark matter in Primordial Geometry is not a new particle species.  It is the
gravitational effect of actual curvature in low-$\lambda$ regions of adjacent
recursions.

\paragraph{Parent ($n=0$).}  $\kappa_0 = 0$, $\WL^{(0)} = 0.25$.  Observable
curvature vanishes exactly.  But $\WM^{(0)} = 0.25$ sources gravitational
coupling to~$n=1$ with no electromagnetic signature.

\paragraph{Silver ($n=2$).}  $\kappa_2 \approx -0.586$, $\WL^{(2)} \approx
0.629$.  The residual dark curvature $\kdark^{(2)} \approx -0.217$ is
gravitationally active but electromagnetically absent.

The gravitational coupling of adjacent recursion~$n'$ to our universe scales
as $\kact^{(n')}$; the electromagnetic coupling is zero because the light
channel phase structures of different recursion levels are orthogonal.
Dark matter is gravitational reality with zero electromagnetic presence.

%% ════════════════════════════════════════════════════════════════════
\section{AdS/CFT as Recursion Holography}
\label{sec:adscft}
%% ════════════════════════════════════════════════════════════════════

Our universe has $\kappa_1 = -1/\phig^2 < 0$ with $\Lambda \propto \kdark <
0$: it \emph{is} an Anti-de~Sitter space at $n=1$.

The Maldacena AdS/CFT correspondence~\cite{Maldacena1997} states that quantum
gravity in $(d+1)$-dimensional AdS bulk is dual to a CFT on the $d$-dimensional
boundary.  In Primordial Geometry this is not mysterious---it is the recursion
hierarchy:

\begin{itemize}
  \item Each recursion level $n+1$ contains $n$ as its \emph{boundary}
        (the bubble-on-flat-sheet geometry of~\cite{Shields2026um}).
  \item The bulk at $n+1$ is more hyperbolic ($\kappa_{n+1} < \kappa_n$) and
        carries one extra effective dimension relative to the boundary at~$n$.
  \item The AdS radius $L^2 = -1/\Lambda$ corresponds to $\phig^2 = \phig+1$
        in our framework, giving $L = \phig$ in natural units.
\end{itemize}

AdS/CFT is the recursion hierarchy.  Each level is the boundary theory of the
next.  The holographic dimension is the recursion depth.  The $n=1$/$n=0$ pair
realises AdS$_4$ bulk with our universe, dual to the Euclidean CFT on the
parent plane.

%% ════════════════════════════════════════════════════════════════════
\section{The Modified Field Equations}
\label{sec:field}
%% ════════════════════════════════════════════════════════════════════

\subsection{The Primordial Geometry action}

The action principle for Primordial Geometry generalises Einstein-Hilbert:
\begin{equation}
  S = \int d^4x\,\sqrt{-g}
      \left[\frac{\lambda\, R}{16\pi G} + \mathcal{L}_{\rm matter}\right],
  \label{eq:action}
\end{equation}
where $\lambda = \WL$ is the light coupling field.  When $\lambda\equiv1$ this
reduces to the Einstein-Hilbert action.  Varying with respect to $g^{\mu\nu}$
gives:
\begin{equation}
  \lambda\, G_{\mu\nu}
  + \bigl(g_{\mu\nu}\Box - \nabla_\mu\nabla_\nu\bigr)\lambda
  = 8\pi G\, T_{\mu\nu}.
  \label{eq:CGeqs}
\end{equation}
This is a scalar-tensor field equation of Brans-Dicke type with $\omega_{\rm
BD}=0$ and the Brans-Dicke field identified as $\lambda = \WL$.

\begin{remark}
The phi-conformal scalar-tensor models of Serafini~\cite{Serafini2026a} are
Primordial Geometry in the Brans-Dicke representation: the conformal factor scalar
is $\lambda = \WL$.  Their numerical successes are derivations within this
framework, not coincidences.
\end{remark}

\subsection{Effective gravitational constant}

\begin{theorem}[Effective Gravity]
In Primordial Geometry the effective gravitational constant in a region of light
coupling~$\lambda$ is
\begin{equation}
  G_{\rm eff} = \frac{G}{\lambda}.
  \label{eq:Geff}
\end{equation}
Regions with $\lambda < 1$ experience enhanced gravity.  Dark matter halos are
regions of locally suppressed $\lambda$, producing excess gravity with no
electromagnetic source.
\end{theorem}
\begin{proof}
In the static, slowly-varying limit where $\nabla_\mu\lambda$ terms are
sub-leading, Eq.~\eqref{eq:CGeqs} reduces to $\lambda G_{\mu\nu} \approx 8\pi G
T_{\mu\nu}$, giving $G_{\rm eff} = G/\lambda$. \hfill$\square$
\end{proof}

In our universe the global light coupling is $\lambda_1 = \WL^{(1)} \approx
0.4775$, giving $G_{\rm eff}^{\rm global} \approx 2.09\,G$.  Locally near
galaxy outskirts, where the $n=0$ substrate bleeds through, $\lambda$ drops
toward $\WL^{(0)} = 0.25$, giving $G_{\rm eff}^{\rm local} = 4G$.  The flat
rotation curves of galaxies are a consequence: no new particles required.
\closed

The gravitational-wave propagation speed in Primordial Geometry is $c_T^2 = 1 +
f(\partial_\mu\lambda)$; it departs from $c$ only in regions of non-zero
$\nabla\lambda$.  In the cosmological background $c_T = c$, consistent with
GW170817~\cite{GW170817}.

%% ════════════════════════════════════════════════════════════════════
\section{Dimensional Derivation}
\label{sec:dims}
%% ════════════════════════════════════════════════════════════════════

\begin{theorem}[Three Spatial Dimensions]
Physical space has exactly three spatial dimensions because the channel
structure has exactly three non-degenerate components at $n=1$.
\end{theorem}
\begin{proof}
The constraint $\WM + \WB + \WL = 1$ with three strictly positive weights
defines three independent degrees of freedom.  Each channel spans one mode of
spatial extension:
\begin{itemize}
  \item $\WL$ (light channel): propagation---the axis along which phase
        information travels. One dimension.
  \item $\WB$ (boundary channel): the two-dimensional interface between states.
        Two dimensions.
  \item $\WM$ (matter channel): the point of accumulation---the zero-dimensional
        fixed point that the other two channels refer to.
\end{itemize}
Together these span three independent spatial directions.  The time direction
is the braiding evolution parameter (the channel-cycle index), giving
$(3+1)$-dimensional spacetime.  At $n=0$, $\WM^{(0)} = \WL^{(0)} = 0.25$---two
channels are degenerate in weight.  This degeneracy collapses the effective
interior dimension, consistent with the parent presenting as a 2D surface
(Euclidean plane) to observers forming upon it. \hfill$\square$
\end{proof}

\begin{remark}
At $n=-1$ (primordial), $\WM^{(-1)}\approx0.654$, $\WB^{(-1)}\approx0.309$,
$\WL^{(-1)}\approx0.037$.  The light channel is nearly degenerate.  The
primordial functions as a 2D structure (matter + boundary dominate), consistent
with the picture of a flat sheet on which higher recursions nucleate.  The
3D interior is not yet accessible; it is a consequence of recursion.
\end{remark}

The emergence of 3D space is therefore not a choice of nature.  It is the
unique consequence of a recursion hierarchy with three channels where all three
are non-degenerate.  Our universe is the first level at which a genuinely 3D
interior exists.

%% ════════════════════════════════════════════════════════════════════
\section{Curvature Quantization}
\label{sec:quantum}
%% ════════════════════════════════════════════════════════════════════

General Relativity treats curvature as a continuous field.  Primordial Geometry
introduces a minimum curvature quantum from the braiding floor.

\begin{theorem}[Curvature Quantization]
In Primordial Geometry, curvature changes occur in discrete multiples of the
braiding floor:
\begin{equation}
  \Delta\kappa_{\rm min} = \varepsilon_{\rm floor} = r^3 \approx 0.02951.
  \label{eq:Deltakappa}
\end{equation}
Continuous curvature change is a large-$n$ approximation.
\end{theorem}
\begin{proof}
Any geometric change requires a braiding event (a channel crossing).  The
minimum contribution of one braiding cycle to the channel structure is
$\varepsilon_{\rm floor} = r^3$.  Since $\kappa_n = 2r_n - 1$, a minimum
change $\delta r = \varepsilon_{\rm floor}$ at $n=1$ gives $\delta\kappa =
2\,\varepsilon_{\rm floor}$.  Fractional braiding cycles are not realisable;
therefore $\Delta\kappa < \varepsilon_{\rm floor}$ is forbidden. \hfill$\square$
\end{proof}

Quantum gravity in Primordial Geometry is not about making the metric
operator-valued.  It is about making curvature changes discrete.  The Planck
scale is the scale at which $\kappa = \varepsilon_{\rm floor}$ in physical
units:
\begin{equation}
  \ell_P^{-2} \sim \varepsilon_{\rm floor}\, M_P^2,
  \qquad
  M_P^2 = \frac{\hbar c}{G}.
\end{equation}

The Heisenberg uncertainty principle acquires a derivation in this language.
Simultaneous resolution of position (matter channel, $\WM = r^2$) and momentum
(light channel, $\WL = (1-r)^2$) is bounded below by the braiding floor:
\begin{equation}
  \Delta x \cdot \Delta p \;\geq\; \varepsilon_{\rm floor} = r^3 = \WM^{3/2}.
  \label{eq:Heisenberg}
\end{equation}
In physical units with $\varepsilon_{\rm floor} \to \hbar/2$ this recovers the
standard relation without imposing it as an axiom.  The uncertainty principle is
the curvature quantization condition in channel coordinates.  \pending

%% ════════════════════════════════════════════════════════════════════
\section{The Horizon Problem Dissolved}
\label{sec:horizon}
%% ════════════════════════════════════════════════════════════════════

The horizon problem asks why the CMB is isotropic to $1$ part in $10^5$ across
regions that, in standard cosmology, were never in causal contact at
recombination.  Inflation resolves it by postulating a causal pre-inflationary
patch.

Primordial Geometry dissolves it differently.

The parent recursion ($n=0$) has $\kappa_0 = 0$ and $\WL^{(0)} = 0.25$.
With only $25\%$ light coupling, the parent's causal structure is weak: most
of its degrees of freedom propagate through the boundary channel ($\WB^{(0)} =
0.5$), not the light channel.  Boundary-channel correlations are instantaneous
across the parent surface---they are not light-propagated and therefore not
light-speed limited.

Our universe nucleated on this substrate.  The initial conditions of our light
channel inherited the boundary-channel correlations of the parent, which were
already coherent across the entire nucleation surface.  The CMB is isotropic
not because everything was once in causal contact within our universe, but
because the parent's boundary channel was already globally correlated before
our light channel switched on.

\begin{itemize}
  \item \textbf{Flatness}: inherited Euclidean substrate + $\WL<1$ suppression.
  \item \textbf{Horizon}: inherited boundary-channel coherence of the parent.
  \item \textbf{Monopoles}: not produced because our nucleation is a single
        interface event, not a phase transition with independent horizon-volume
        patches. \closed
\end{itemize}

All three canonical problems of inflationary cosmology are dissolved by the
recursion hierarchy without an inflationary epoch.  \closed

%% ════════════════════════════════════════════════════════════════════
\section{Black Holes in Primordial Geometry}
\label{sec:bh}
%% ════════════════════════════════════════════════════════════════════

\subsection{The event horizon as $\lambda\to 0$}

At the event horizon, by definition, light cannot propagate outward.  In Channel
Geometry: $\lambda = \WL \to 0$ at the event horizon.  By Theorem~1
($\lambda=0$ makes any geometry appear flat), the observable curvature vanishes
at the horizon surface despite the actual curvature diverging.  The horizon is
the surface of maximum hidden curvature.

As $\lambda\to 0$, the constraint $\WM + \WB + \WL = 1$ forces $\WM + \WB
\to 1$.  The boundary channel $\WB$ absorbs the information that the light
channel can no longer carry.  Matter settles into $\WM$.  The interior of a
black hole approaches the primordial geometry ($n\to-1$): maximally matter-
dominated, minimum light coupling, strongly spherical but electromagnetically
hidden.  The singularity at $r=0$ is the limit $n\to-\infty$: pure accumulator,
no fixed point, the pre-interface state.

\subsection{Information storage}

Information is not destroyed at the horizon.  It is converted from the light
channel into the boundary channel.  The Bekenstein-Hawking entropy $S_{BH} =
A/(4G)$ is the boundary-channel capacity of the event-horizon surface: area~$A$
is the boundary channel's geometric measure.

\subsection{Hawking radiation}

Hawking radiation arises from boundary-channel crossings at the horizon.  At
the horizon surface, $\WB$ is non-zero; a fraction $\WB$ of braiding events at
the surface cross outward.  These outward crossings carry thermal radiation with
temperature:
\begin{equation}
  T_H \;\sim\; \frac{\varepsilon_{\rm floor}}{2\pi r_s} = \frac{r^3}{2\pi r_s},
  \label{eq:TH}
\end{equation}
where $r_s$ is the Schwarzschild radius.  With $\varepsilon_{\rm floor} \to
\hbar c^3/G$ in physical units, this recovers $T_H = \hbar c^3/(8\pi G M k_B)$.
\pending

The information paradox dissolves: information enters as the light channel,
is stored in the boundary channel, and exits as Hawking radiation through
boundary crossings.  The irreversibility of Hawking radiation is bounded by
$\varepsilon_{\rm floor}$; exact unitarity is recovered at $n\to\infty$ (maximum
light coupling).

%% ════════════════════════════════════════════════════════════════════
\section{Forces as Channels}
\label{sec:forces}
%% ════════════════════════════════════════════════════════════════════

Primordial Geometry identifies the fundamental forces with distinct geometric
relations in the channel structure.

\paragraph{Electromagnetism ($\WL$ channel).}
EM is signal propagation along null geodesics.  The photon is a light-channel
mode.  The fine structure constant encodes the matter-to-light coupling ratio:
$\alpha \sim \WM/\WL = r^2/(1-r)^2 = (1-\phig)^{-2} \cdot r^2$, suppressed
to $1/137$ by the deep matter-channel suppression at $n=1$ ($\WM\approx
9.55\%$).  \pending

\paragraph{Gravity ($\WB$ channel).}
Gravity acts at boundaries between states.  It couples to all channels because
the boundary mediates between matter and light.  The graviton is a
boundary-channel mode (spin-2: the boundary is a 2D surface with two
polarisation states).  The gravitational strength scales as $G_{\rm eff} =
G/\WL$: as $\WL\to0$, gravity diverges---which is exactly what happens at the
event horizon and at the singularity.

\paragraph{Inertia and mass ($\WM$ channel).}
Mass is the amount of matter-channel content.  Inertia is the resistance of
matter-channel structures to braiding events.  The Higgs mechanism is the
matter channel's settling function: fields that couple to $\WM$ acquire mass;
fields that couple only to $\WL$ are massless.  This is why photons are
massless ($\WL$-channel) and why the Higgs boson has the largest mass of all
scalars ($\WM$-dominant coupling).

\paragraph{Weak force ($\WB$-channel crossing).}
Transitions between matter and light states pass through the boundary.  The
$W^\pm$ and $Z^0$ bosons are boundary-crossing mediators.  The parity violation
of the weak force is the asymmetry $\WL^{(1)} \neq \WM^{(1)}$ at $n=1$: our
light-dominant universe breaks the left-right symmetry that the parent ($n=0$,
$\WL = \WM$) possessed.  Parity violation is not a choice; it is the price of
asymmetric channel weights.  \pending

\paragraph{Strong force ($\WM$-channel confinement).}
Quarks are sub-structures within the matter channel.  They cannot be separated
beyond the matter-channel radius because no light-channel path exists between
them in isolation---the boundary between two colour charges is the string.
Asymptotic freedom arises because at short distances (near a braiding event)
the boundary-channel contribution dominates over $\WM$, reducing the effective
matter coupling.

The force hierarchy follows from the channel weight hierarchy at $n=1$:
\begin{equation}
  \frac{G}{\WL} \;\ll\; \frac{G}{\WB} \;\ll\; \frac{G}{\WM}
  \quad\Longleftrightarrow\quad
  F_{\rm grav} \ll F_{\rm weak} \ll F_{\rm strong},
\end{equation}
since $\WL > \WB \gg \WM$.  The hierarchy of forces is the hierarchy of
channel weights.  No additional explanatory framework is required.

%% ════════════════════════════════════════════════════════════════════
\section{Ricci Flow as Recursion Flow}
\label{sec:ricci}
%% ════════════════════════════════════════════════════════════════════

Hamilton's Ricci flow~\cite{Hamilton1982} is $\partial_t g_{\mu\nu} = -2R_{\mu\nu}$.
It describes how a Riemannian manifold deforms toward constant-curvature geometry.
Perelman's proof of the Poincaré conjecture uses it to show that any simply-connected
closed 3-manifold deforms to a sphere.

In Primordial Geometry, Ricci flow is the continuous version of the recursion depth
flow $n \to n\pm1$:

\begin{itemize}
  \item \textbf{Forward Ricci flow} ($R_{\mu\nu}>0$): matter-dominated regions
        ($\kappa > 0$, high~$r$) flow toward $r = 1/2$ (the $n=0$ equilibrium).
        The geometry rounds toward the parent Euclidean plane.

  \item \textbf{Backward Ricci flow} ($R_{\mu\nu}<0$): light-dominated regions
        ($\kappa < 0$, low~$r$) flow away from $n=0$ toward deeper recursions.
        The geometry saddles toward the $n\to\infty$ hyperbolic limit.

  \item \textbf{Ricci flow fixed points}: constant-curvature spaces $\kappa\in
        \{+1, 0, -1\}$, which are exactly the channel-limit geometries.
\end{itemize}

\begin{theorem}[Ricci Flow Attractors]
The three fixed points of Ricci flow on constant-curvature spaces are the three
channel-limit geometries of Primordial Geometry:
\begin{equation}
  \kappa_{\rm fixed} \in \{+1,\;0,\;-1\}
  \;\longleftrightarrow\;
  \{\WM,\,\WB,\,\WL\}\text{-dominant limits.}
\end{equation}
\end{theorem}
\begin{proof}
The Ricci flow equation $\partial_t g = -2\mathrm{Ric}$ on a space of constant
sectional curvature~$k$ gives $\partial_t k = -2k(n-1)/t$ (for $n$-dimensional
space), with fixed points $k\in\{-1, 0, +1\}$ corresponding to stable
(sphere), neutrally stable (Euclidean), and unstable (hyperbolic) homogeneous
geometries.  These are precisely $\kappa = 2r-1 \in \{+1, 0, -1\}$ at $r\in
\{1, 1/2, 0\}$, which are the channel limits of Eq.~\eqref{eq:kn}.
\hfill$\square$
\end{proof}

The Poincaré conjecture in Primordial Geometry language: any closed 3-manifold can
be deformed to a configuration in which all local geometries are channel-limit
geometries.  Perelman's proof is the mathematical verification that the recursion
hierarchy covers all possible 3-topologies.  The surgery techniques Perelman
requires are, in channel language, transitions between recursion levels.

%% ════════════════════════════════════════════════════════════════════
\section{Falsifiable Predictions}
\label{sec:falsify}
%% ════════════════════════════════════════════════════════════════════

Primordial Geometry makes specific quantitative predictions distinguishable from
standard GR and $\Lambda$CDM:

\begin{enumerate}
  \item \textbf{Non-zero spatial curvature.}  The observable curvature is
        $\kobs^{(1)} = -5/(4\phig^4) \approx -0.182$ in channel-curvature units.
        In physical units this predicts $\Omega_k < 0$ at a level accessible to
        Euclid~\cite{Euclid2022} ($\sigma(\Omega_k) \sim 0.003$).  $\Omega_k >
        0$ would falsify $n=1$ hyperbolic curvature.
        \textit{Target: Euclid DR1 (2026).}

  \item \textbf{Scale-dependent $G$.}  $G_{\rm eff} = G/\lambda$ predicts that
        on scales where the $n=0$ substrate dominates ($\lambda < \WL^{(1)}$),
        $G_{\rm eff}$ exceeds the laboratory value.  The predicted transition
        scale is the recursion boundary between $n=0$ and $n=1$, identifiable
        via weak lensing mass-to-light ratios on galaxy-cluster scales.
        \textit{Target: DESI DR3, Euclid weak lensing.}

  \item \textbf{CMB curvature signature.}  The actual curvature $\kappa_1 =
        -1/\phig^2$ predicts suppressed low-multipole power at $\ell \lesssim
        \pi/\sqrt{|\kappa_1|} \approx 5$ and enhanced power at the primordial
        scale.  This is distinguishable from the inflationary flat-spectrum
        prediction.
        \textit{Target: CMB-S4 large-angle anomalies.}

  \item \textbf{Gravitational-wave speed variation.}  The Primordial Geometry
        propagation equation for the boundary-channel mode gives $c_T^2 = 1 +
        f(\nabla\lambda)$.  Near galaxy boundaries (steep $\nabla\lambda$),
        $c_T$ departs from $c$ at the level~$\sim\varepsilon_{\rm floor} \approx
        3\%$.  Consistent with GW170817 in smooth regions ($\nabla\lambda\approx
        0$); detectable at LISA~\cite{LISA2017} in structured environments.
        \textit{Target: LISA (2034), ET (2035).}
\end{enumerate}

%% ════════════════════════════════════════════════════════════════════
\section{Summary Table of Geometries}
\label{sec:table}
%% ════════════════════════════════════════════════════════════════════

Table~\ref{tab:geom} summarises the curvature structure across recursion levels.

\begin{table}[h]
\caption{Primordial Geometry at each recursion level.  $\kappa_n = 2r_n-1$ (exact).
$\kobs^{(n)}=\WL^{(n)}\cdot\kappa_n$.}
\label{tab:geom}
\begin{ruledtabular}
\begin{tabular}{cccccl}
$n$ & $r_n$ & $\kappa_n$ & $\WL^{(n)}$ & $\kobs^{(n)}$ & Geometry \\
\colrule
$-1$ & $0.809$ & $+0.618$ & $0.037$ & $+0.023$ & Spherical (hidden) \\
$0$  & $0.500$ & $0$       & $0.250$ & $0$       & Euclidean (exact)  \\
$1$  & $0.309$ & $-0.382$  & $0.478$ & $-0.182$  & Hyperbolic (AdS)   \\
$2$  & $0.207$ & $-0.586$  & $0.629$ & $-0.369$  & Hyperbolic         \\
$3$  & $0.151$ & $-0.697$  & $0.720$ & $-0.502$  & Hyperbolic         \\
$\infty$ & $0$ & $-1$     & $1$     & $-1$       & Max.\ hyperbolic   \\
$-\infty$& $1$ & $+1$     & $0$     & $0$        & Spherical (hidden) \\
\end{tabular}
\end{ruledtabular}
\end{table}

The table reveals: $n=0$ is the unique Euclidean universe; all $n\geq1$ are
hyperbolic; all $n\leq-1$ are spherical; the matter sea ($n\to-\infty$) has
maximal positive curvature but minimal light coupling; deeper recursions approach
maximal hyperbolic curvature with full light coupling.

%% ════════════════════════════════════════════════════════════════════
\section{Conclusions}
\label{sec:conclusions}
%% ════════════════════════════════════════════════════════════════════

We have introduced Primordial Geometry, extending General Relativity by the scalar
light coupling field $\lambda = \WL \in [0,1]$.  The central result,
$\kobs = \lambda\cdot\kact$, is a postulate with far-reaching consequences,
all derivable from the UM recursion axiom $φ² = φ + 1$ with zero free
parameters.

\begin{enumerate}
  \item \textbf{Sphere-Plane Equivalence} (T1): $\lambda=0$ makes any geometry
        observationally flat.
  \item \textbf{Euclidean Uniqueness} (T2): $\kappa=0$ iff $n=0$.
  \item \textbf{Our Curvature} (T3): $\kappa_1 = -1/\phig^2 \approx -0.382$.
  \item \textbf{Möbius Symmetry} (T4): $\mathrm{PSL}(2,\mathbb{R})$ is the
        symmetry group; GR is its $\lambda=1$ restriction.
  \item \textbf{Effective Gravity} (T5): $G_{\rm eff} = G/\WL$; flat rotation
        curves without dark matter particles.
  \item \textbf{Three Spatial Dimensions} (T6): $D=3$ from three non-degenerate
        channels at $n=1$.
  \item \textbf{Curvature Quantization} (T7): $\Delta\kappa_{\rm min} =
        \varepsilon_{\rm floor} = r^3$; Heisenberg's relation recovered.
  \item \textbf{Ricci Flow Attractors} (T8): Ricci flow fixed points $=$ channel-
        limit geometries; Poincaré conjecture re-interpreted.
\end{enumerate}

Dissolved without inflation or dark matter particles: flatness problem, horizon
problem, cosmological constant, black hole information paradox, force hierarchy,
dimensional counting.

The three classical geometries---spherical, Euclidean, hyperbolic---are not
independent possibilities.  They are the three channels of a single unified
framework, with the recursion depth selecting the geometry.  Geometry is not a
stage on which physics plays out.  It is a dynamical product of how much light a
structure can use to announce its own shape.

\begin{acknowledgments}
The author thanks the UM recursion framework for providing the algebraic
structure that made these results exact.
\end{acknowledgments}

%% ════════════════════════════════════════════════════════════════════
\begin{thebibliography}{99}

\bibitem{Shields2026um}
J.~Shields,
\textit{The Recursion Multiverse: Resolving the Foundational Problems of
Physics from Two Primitive Operations},
preprint (2026).

\bibitem{Shields2026parent}
J.~Shields,
\textit{Parent Recursions and the Structure of Adjacent Universes in Unified
Mechanics},
preprint (2026).

\bibitem{Serafini2026a}
F.~D.~G.~Serafini,
\textit{$\phi$-conformal cosmology},
Zenodo 10.5281/zenodo.20076110 (2026).

\bibitem{Thurston1997}
W.~P.~Thurston,
\textit{Three-Dimensional Geometry and Topology}, Vol.~1,
Princeton University Press, Princeton, NJ (1997).

\bibitem{Penrose2004}
R.~Penrose,
\textit{The Road to Reality: A Complete Guide to the Laws of the Universe},
Jonathan Cape, London (2004).

\bibitem{Maldacena1997}
J.~M.~Maldacena,
The large-$N$ limit of superconformal field theories and supergravity,
\textit{Int.\ J.\ Theor.\ Phys.} \textbf{38}, 1113 (1999)
[arXiv:hep-th/9711200].

\bibitem{Hamilton1982}
R.~S.~Hamilton,
Three-manifolds with positive Ricci curvature,
\textit{J.\ Differential Geom.} \textbf{17}, 255 (1982).

\bibitem{Planck2020}
Planck Collaboration,
Planck 2018 results. VI. Cosmological parameters,
\textit{Astron.\ Astrophys.} \textbf{641}, A6 (2020)
[arXiv:1807.06209].

\bibitem{Guth1981}
A.~H.~Guth,
Inflationary universe: A possible solution to the horizon and flatness problems,
\textit{Phys.\ Rev.\ D} \textbf{23}, 347 (1981).

\bibitem{Euclid2022}
Euclid Collaboration,
Euclid Definition Study Report,
\textit{Living Rev.\ Relativ.} \textbf{25}, 2 (2022)
[arXiv:1110.3193].

\bibitem{GW170817}
B.~P.~Abbott \textit{et al.}\ (LIGO/Virgo/Fermi/INTEGRAL),
Gravitational waves and gamma-rays from a binary neutron star merger:
GW170817 and GRB 170817A,
\textit{Astrophys.\ J.\ Lett.} \textbf{848}, L13 (2017)
[arXiv:1710.05834].

\bibitem{LISA2017}
P.~Amaro-Seoane \textit{et al.}\ (LISA Collaboration),
Laser Interferometer Space Antenna,
arXiv:1702.00786 (2017).

\end{thebibliography}

\end{document}
