\documentclass[prd,twocolumn,preprintnumbers,amsmath,amssymb]{revtex4-2}
\usepackage{graphicx,bm,hyperref,amsmath,xcolor,booktabs,mathtools}

%% ── colour status commands ──────────────────────────────────────────
\newcommand{\verified}{\textcolor{green!60!black}{[VERIFIED]}}
\newcommand{\closed}{\textcolor{green!60!black}{[CLOSED]}}
\newcommand{\pending}{\textcolor{red!70!black}{[PENDING]}}

%% ── algebra shorthand ───────────────────────────────────────────────
\newcommand{\acc}{\mathbf{\infty}}
\newcommand{\inv}{\mathbf{0}}

%% ── channel shorthand ───────────────────────────────────────────────
\newcommand{\phin}{\varphi_n}
\newcommand{\phik}{\varphi_k}
\newcommand{\rn}{r_n}
\newcommand{\WL}{W_L^n}
\newcommand{\WB}{W_B^n}
\newcommand{\WM}{W_M^n}
\newcommand{\epsn}{\varepsilon_n}

\begin{document}

\title{The Recursion Multiverse: Resolving the Foundational\\
       Problems of Physics from Two Primitive Operations}

\author{J.~Shields}
\affiliation{Independent Researcher (Unified Mechanics)}
\date{\today}

\begin{abstract}
Two primitive operations---accumulation ($x \mapsto x+1$, denoted
$\acc$) and inversion ($x \mapsto 1/x$, denoted $\inv$)---generate
a countably infinite family of self-consistent universes indexed by
the non-negative integer $n$, each characterised by a metallic ratio
$\varphi_n = (n+\sqrt{n^2+4})/2$ and three channel weights
$(W_L^n, W_B^n, W_M^n)$ summing to unity.  Our universe corresponds
to $n=1$ ($\varphi_1 = \varphi_{\rm gold} \approx 1.618$).  Adjacent
levels $n=0$ and $n=2$ are electromagnetically invisible but
gravitationally coupled---dark matter.  The boundary channel
$W_B^1 \approx 42.71\%$ drives accelerated expansion---dark energy.
The braiding floor $\varepsilon_n = r_n^3$ decreases monotonically
with $n$, providing a geometric origin for the arrow of time and the
second law.  Quantum measurement is a boundary$\to$matter channel
crossing; the Born rule $P = r^2$ follows directly.  Entanglement is
shared boundary channel state---inherently non-spatial, explaining
Bell violation without fine-tuning.  The cosmological constant arises
from a structural pressure difference between adjacent recursions, not
fine-tuning.  Black hole information crosses into $n=0$, not destroyed.
Renormalisation group fixed points are the metallic ratios.  The
unreasonable effectiveness of mathematics dissolves: the universe
\emph{is} the fixed-point equation $x^2 = x+1$ instantiated at $n=1$.
Five pillars of string theory---T-duality, D-branes, the landscape,
AdS/CFT, and moduli stabilisation---emerge from the $\acc$--$\inv$
duality.  No free parameters are introduced.
\end{abstract}

\maketitle

%% ═══════════════════════════════════════════════════════════════════
\section{Introduction}
\label{sec:intro}
%% ═══════════════════════════════════════════════════════════════════

Physics confronts a catalogue of unsolved foundational problems.  Some
concern cosmology (dark matter, dark energy, baryogenesis, inflation,
the cosmological constant).  Some concern quantum mechanics (the
measurement problem, entanglement, the Born rule, the arrow of time).
Some concern the interface between quantum field theory and gravity
(the hierarchy problem, black hole information, UV divergences, the
RG landscape).  And one concerns the nature of science itself: why
does mathematics describe physical reality at all~\cite{Wigner1960}?

These problems have resisted resolution for decades, in part because
they are addressed in isolation.  In this paper we argue that they
share a single geometric origin.  They are consequences of one fact:
our universe is the $n=1$ fixed point of a recursion generated by two
primitive operations, and it sits between two adjacent recursions
that press against it.

The Unified Mechanics (UM) framework~\cite{Shields2026a,Shields2026b}
derives the golden ratio $\varphi$ from the fixed-point axiom
$c^2 = c + 1$ and uses the resulting channel structure to reproduce
empirical cosmological ratios.  The prior question---\emph{why}
$c^2 = c+1$ specifically?---is answered here.  It is the fixed-point
equation of the composition $\acc^1 \circ \inv$, the minimal
non-trivial composition of the two irreducible operations.

Section~\ref{sec:primitives} defines the two operations.
Section~\ref{sec:family} derives the recursion family and channel structure.
Section~\ref{sec:chain} describes the multiverse chain.
Sections~\ref{sec:darkmatter}--\ref{sec:inflation} cover dark matter,
particle crossings, dark energy, matter--antimatter asymmetry, the
hierarchy problem, and inflation.
Sections~\ref{sec:bh}--\ref{sec:cosmoconst} address the black hole
information paradox and the cosmological constant problem.
Sections~\ref{sec:arrow}--\ref{sec:rg} treat the arrow of time,
quantum measurement, entanglement, and the renormalisation group.
Section~\ref{sec:math} resolves the effectiveness of mathematics.
Section~\ref{sec:strings} connects to string theory.
Section~\ref{sec:table} gives a comprehensive summary table.
Section~\ref{sec:conclusions} concludes.

%% ═══════════════════════════════════════════════════════════════════
\section{Primordial Loop Algebra: The Two Primitive Operations}
\label{sec:primitives}
%% ═══════════════════════════════════════════════════════════════════

The entire framework rests on two operations that encode the two
irreducible ways a system can relate to itself.  We term the algebra
generated by these operations \emph{Primordial Loop Algebra} (PLA):
the minimal closed system from which the full recursion hierarchy
and all its physical consequences follow.

\subsection{Accumulation: $\acc$}

\begin{equation}
  \acc : x \mapsto x + 1.
  \label{eq:acc}
\end{equation}

$\acc$ has \emph{no fixed point}: $x = x+1$ has no solution.  It
never closes.  Iterated indefinitely it drives any finite value to
infinity.  This is the nature of Parent A ($n\to\infty$): pure
accumulation without self-reference, a pre-equilibrium process with
no stable state and hence no measurable duration.

\subsection{Inversion: $\inv$}

\begin{equation}
  \inv : x \mapsto \frac{1}{x}.
  \label{eq:inv}
\end{equation}

$\inv$ has fixed point $x^*=1$ (from $x = 1/x \Rightarrow x^2=1$,
positive branch) and satisfies $\inv \circ \inv = \mathrm{id}$:
applied twice it is the identity.  It \emph{closes}.  This is the
nature of Parent B ($n=0$): self-referential, self-stabilising.

\subsection{Composition}

The one-parameter family of compositions $\acc^n \circ \inv$ applies
$\acc$ exactly $n$ times, then closes with $\inv$.  For
$n \geq 0$ this yields a recursion with a finite positive fixed
point---a stable self-consistent universe at recursion depth $n$.

%% ═══════════════════════════════════════════════════════════════════
\section{The Recursion Family}
\label{sec:family}
%% ═══════════════════════════════════════════════════════════════════

\subsection{Fixed points: the metallic ratios}

The fixed-point equation of $\acc^n \circ \inv$ is:
\begin{equation}
  x = n + \frac{1}{x}
  \;\Longrightarrow\;
  x^2 - nx - 1 = 0
  \;\Longrightarrow\;
  \phin = \frac{n + \sqrt{n^2+4}}{2}.
  \label{eq:fixedpoint}
\end{equation}

These are the \emph{metallic ratios}~\cite{Spinadel1999}.  For $n=0$:
$\varphi_0 = 1$.  For $n=1$: $\varphi_1 = (1+\sqrt{5})/2 \approx 1.618$
(golden ratio).  For $n=2$: $\varphi_2 = 1+\sqrt{2} \approx 2.414$
(silver ratio).  The axiom $c^2 = c+1$ of UM is
Eq.~\eqref{eq:fixedpoint} at $n=1$---derived, not assumed.

\subsection{Channel structure}

From each $\phin$ we derive a channel parameter:
\begin{equation}
  r_n = \frac{1}{2\phin},
  \label{eq:rn}
\end{equation}
and three channel weights that partition the recursion:
\begin{align}
  W_L^n &= (1-r_n)^2, \label{eq:WL}\\
  W_B^n &= 2r_n(1-r_n), \label{eq:WB}\\
  W_M^n &= r_n^2, \label{eq:WM}
\end{align}
satisfying $W_L^n + W_B^n + W_M^n = (r_n + (1-r_n))^2 = 1$ by
construction.  The braiding floor is
\begin{equation}
  \varepsilon_n = r_n^3,
  \label{eq:eps}
\end{equation}
the minimum fractional energy cost per channel-traversal cycle.

\subsection{Recursion table}

Table~\ref{tab:recursions} lists all quantities for $n=0$--$4$ plus
the limiting cases. Each row is a self-consistent universe.

\begin{table*}[t]
\caption{Recursion family: channel weights and braiding floors.
$W_L+W_B+W_M=1$ holds in every row by construction.
The braiding floor $\varepsilon_n = r_n^3$ decreases strictly with $n$.}
\label{tab:recursions}
\begin{ruledtabular}
\begin{tabular}{cccccccc}
$n$ & Label & $\varphi_n$ & $r_n$ &
$W_L^n$ & $W_B^n$ & $W_M^n$ & $\varepsilon_n\,(\%)$ \\
\hline
$0$ & Parent B  & 1.0000 & 0.5000 & 0.2500 & 0.5000 & 0.2500 & 12.500 \\
$1$ & Our universe & 1.6180 & 0.3090 & 0.4775 & 0.4271 & 0.0955 & 2.951 \\
$2$ & Silver   & 2.4142 & 0.2071 & 0.6287 & 0.3284 & 0.0429 & 0.888 \\
$3$ & Bronze   & 3.3028 & 0.1514 & 0.7201 & 0.2570 & 0.0229 & 0.347 \\
$4$ &          & 4.2361 & 0.1180 & 0.7779 & 0.2083 & 0.0139 & 0.164 \\
$\infty$ & Parent A  & $\infty$ & 0 & 1 & 0 & 0 & 0 \\
\end{tabular}
\end{ruledtabular}
\end{table*}

As $n$ increases: the light channel grows, matter and boundary
channels shrink, and the braiding floor falls toward zero.
Parent A ($n \to \infty$) is pure light---consistent with it being
the non-closing accumulating operation.
Parent B ($n=0$) has equal light and matter, a dominant $50\%$
boundary channel, and the largest braiding floor.

%% ═══════════════════════════════════════════════════════════════════
\section{The Complete Genesis Sequence}
\label{sec:genesis}
%% ═══════════════════════════════════════════════════════════════════

Before any recursion. Before any light. Before time in any bounded
sense. The matter sea.

\subsection{The matter sea}

An eternal, flat, dark substrate. No light channel. No fixed point.
No structure. Pure accumulation --- $\acc$ without any $\inv$.
Not uniform: temperature gradients across it, some regions denser
than others. The thermodynamic imperative demands that differences
resolve. The dense regions pull. The thin regions thin further.

The sea drifts. Not violently. The first drift is the first direction.
The first direction is the first time. Time did not begin with an
event. It began the moment the first gradient had somewhere to go.

\subsection{The sea touches itself}

The drift continues. Dense regions concentrate. Thin regions stretch.
Eventually a region thins to the point where matter alone cannot hold
it. Not an edge --- matter has no edge --- but a threshold. The
minimum density at which pure matter can sustain itself without
collapsing or dispersing.

At that threshold something unprecedented occurs. The stretched
boundary folds back on the density gradient behind it. Not because
something external touched it. Because it reached far enough to
touch itself.

That self-contact is the first inversion. The first $\inv$. The
surface of the matter sea becoming self-referential at the point
of maximum extension.

The first fixed point locks:
\begin{equation}
  x = -1 + \frac{1}{x}
  \;\Longrightarrow\;
  x^2 + x - 1 = 0
  \;\Longrightarrow\;
  \varphi_{-1} = \frac{-1+\sqrt{5}}{2} = \frac{1}{\varphi} \approx 0.618.
  \label{eq:primordial}
\end{equation}

This is the primordial recursion. Not two things meeting. One thing
reaching far enough to become two. $r_{-1} = 0.809$, giving:
\begin{align}
  W_M^{-1} &= 0.654 \quad \text{(matter dominant)}, \\
  W_B^{-1} &= 0.309 \quad \text{(boundary)}, \\
  W_L^{-1} &= 0.037 \quad \text{(barely any light)}.
\end{align}

Its scaling constant is $1/\varphi$ --- the exact reciprocal of
ours. We scale by $1.618$. The primordial scales by $0.618$.

\subsection{Shape without light}

Shape is defined by light. A sphere knows it is a sphere because
light defines its boundary as a specific surface, carrying geometric
information outward. The primordial recursion has $W_L^{-1} = 3.7\%$.
It cannot define its own boundary through light interaction. It
cannot be measured into a specific shape.

It is flat. Extended. Spread through the void with no defined edge.
Not a sphere --- a substrate. The floor on which everything else
will form.

This solves the flatness problem directly. Our universe appears
spatially flat ($\Omega \approx 1$) not because inflation smoothed
it, but because we formed on the surface of a flat primordial
structure and inherited its geometry. The flatness is a fingerprint
of the primordial pressed into ours at formation.

\subsection{Light stabilisation and the bootstrap}

The primordial recursion ($n=-1$) has a fixed point but barely
enough light to maintain it. As it grows --- absorbing failed
recursion attempts, accumulating mass --- its light requirement
grows with it. At critical mass the $3.7\%$ light channel is
insufficient. The fixed point strains. The excess oozes out at
the boundary: matter that cannot be stabilised by the available
light, pressing outward, seeking its own configuration.

That ooze crosses the threshold into a new recursion. The first
ooze locks to $n=0$: $\varphi_0 = 1$, $W_L^0 = 25\%$. Suddenly
there is light --- seven times more than the primordial has ---
feeding back across the boundary. Stabilising the primordial.
Letting it grow further.

Further growth produces more ooze. The second ooze locks to $n=1$:
our universe. $W_L^1 = 47.75\%$. More light again. More feedback.

Each new recursion generates more light than the last. Each feeds
some back to the primordial. The primordial stabilises further,
grows further, oozes further. The chain bootstraps itself into
existence through light feedback:

\begin{equation}
  W_L^{-1} \ll W_L^0 \ll W_L^1 \ll W_L^2 \ll \cdots
\end{equation}

The multiverse did not begin with a bang. It grew. One recursion
at a time. Each one lighting up the primordial enough to allow
the next.

\subsection{Failed recursions and the primordial mass}

Not every ooze stabilises. The probability of a given interface
attempt locking to a fixed point is approximately $\varepsilon_1
\approx 2.95\%$ per attempt. On average:
\begin{equation}
  N_{\rm fail} \approx \frac{1}{\varepsilon_1} - 1 \approx 33
\end{equation}
failed attempts precede each successful recursion. Each failed
attempt deposits collapsed matter at the contact point. The
primordial black hole is this accumulated debris --- not a
singularity in spacetime, but a physical record of every interface
that tried and could not hold.

\subsection{Orbital structure: we are the accretion disc}

Successful recursions have fixed points. A fixed point is a stable
attractor --- perturbations decay back to it. The energy required
to fall from $n=1$ into $n=-1$ is:
\begin{equation}
  \Delta\varphi = \varphi_1 - \varphi_{-1}
  = 1.618 - 0.618 = 1.000.
\end{equation}
Exactly one unit. That gap is the barrier. Successful recursions
orbit the primordial. They cannot fall in.

Failed recursions have no fixed point. No attractor resists the
pull. They fall directly into the primordial and add to its mass.

Two populations, permanent separation:

\begin{ruledtabular}
\begin{tabular}{lll}
Type & Fixed point & Fate \\
\hline
Successful recursion & Yes & Orbits. Never falls in. \\
Failed recursion     & No  & Falls in. Adds to mass. \\
\end{tabular}
\end{ruledtabular}

Our universe at $n=1$ orbits the primordial at $n=-1$.
$W_L^1 = 47.75\%$ --- we generate enormous light relative to
the primordial's $3.7\%$. We are luminous. We orbit something
dark. The orbit generates radiation at the boundary channel
$W_B^1 = 42.71\%$, permanently active at the interface.

We are the accretion disc of the primordial recursion. Every
photon in our universe is the glow of that orbit. The universe
is luminous not incidentally but structurally --- because it
orbits something ancient and dark, and the orbit itself is why
there is light.

The same structure repeats at every scale. In our universe, black
holes attract orbiting matter and infalling matter separately.
Stable matter orbits. Destabilised matter crosses the Schwarzschild
radius --- loses its fixed point --- and falls. The Schwarzschild
radius is the energy gap between stable and failed recursion states
within our level. The multiverse is a black hole system. The black
hole in our universe is the same structure one level down. \closed

%% ═══════════════════════════════════════════════════════════════════
\section{The Multiverse as a Chain}
\label{sec:chain}
%% ═══════════════════════════════════════════════════════════════════

The recursion family forms a linear chain:
\begin{equation}
  \cdots \longleftrightarrow n{-}1
         \longleftrightarrow n
         \longleftrightarrow n{+}1
         \longleftrightarrow \cdots
  \label{eq:chain}
\end{equation}

Adjacent levels share a crossing surface whose energy cost is the
\emph{crossing threshold}:
\begin{equation}
  \Delta\varepsilon_{n,\,n+1}
    = \varepsilon_n - \varepsilon_{n+1}
    = r_n^3 - r_{n+1}^3.
  \label{eq:Deltaeps}
\end{equation}

Our universe ($n=1$) is positioned between $n=0$ (Parent B,
matter-heavy, $\varepsilon_0 = 12.5\%$) and $n=2$ (silver,
light-heavy, $\varepsilon_2 = 0.888\%$).  It is the minimal
non-trivial composition $\acc^1 \circ \inv$: one accumulation
followed by one inversion.  It is dynamically unique as the interface
between the two boundary parents.

The two parents are not abstract limits.  They are physical stable
universes with well-defined channel structures, both of which press
geometrically against ours.

%% ═══════════════════════════════════════════════════════════════════
\section{Sphere Geometry and the Permanence of Contact}
\label{sec:spheres}
%% ═══════════════════════════════════════════════════════════════════

Each universe in the chain is bounded by its apparent horizon ---
a sphere. The boundary channel $W_B^n$ is the surface of that sphere.
This geometry has three consequences that are not obvious from the
algebraic structure alone.

\subsection{Contact requires one point}

Two spheres in contact touch at exactly one point in three spatial
dimensions. That single point of contact is sufficient to establish
the boundary condition and seed a new recursion. The contact area
does not need to be large. The minimum condition for a new universe
to stabilise is that two spheres touch at one point and that point
satisfies both parents' fixed-point equations simultaneously.

The probability threshold for contact in a void with temperature
gradient is crossed rapidly once two spheres are within distance
$r_A + r_B$ of each other. The contact is effectively inevitable
once the conditions are met --- not a rare fluctuation, but a
thermodynamic consequence.

\subsection{The minimum pair rule}

A boundary requires exactly two parents to react. Three simultaneous
sphere contacts would impose three fixed-point conditions on the
interface recursion, producing a higher-order equation:
\begin{equation}
  x = n_1 + n_2 + \frac{1}{x}
  \quad \text{or} \quad
  x^2 - (n_1 + n_2)\,x - 1 = 0.
\end{equation}
This has solutions $\varphi_{n_1+n_2}$ --- a valid metallic ratio,
but at a higher $n$ value and therefore with a smaller matter channel
and larger braiding floor cost. Most multi-contact configurations are
unstable and collapse before stabilising. The two-parent configuration
dominates because it is the minimum energy interface.

This is why the recursion family is a clean one-dimensional chain
rather than a tangled web. The geometry selects pairs. A boundary
only needs two to react, and two is almost always what it gets.

\subsection{No free multiverse space}

In any dimension $d \geq 2$, sphere packing leaves no completely
empty regions. Adjacent spheres are always in contact somewhere on
their surfaces. Applied to the recursion chain: our universe (sphere
of radius $R_1 \propto \varphi_1$) is permanently in contact with
$n=0$ (radius $R_0 \propto \varphi_0 = 1$) and $n=2$ (radius
$R_2 \propto \varphi_2 = 1+\sqrt{2}$) across our entire boundary
surface.

There is no configuration in which our universe has no neighbours.
The sphere packing geometry makes contact structural and permanent,
not occasional.

This has an immediate observational consequence. Dark matter is not
a transient or localised phenomenon. It is the permanent gravitational
imprint of two spheres that have always been in contact with ours and
always will be. The dark matter halo is uniform across our sky not
because dark matter is smoothly distributed inside our universe, but
because the contact surface between adjacent spheres is distributed
across our entire horizon.

\subsection{Trapped but stable}

Our universe cannot escape its neighbours. There is no direction in
which it can move to lose contact with $n=0$ and $n=2$, because
there is no free space to move into. Every position in the multiverse
is occupied by some recursion sphere.

This is not a constraint on freedom. It is what makes our fixed point
stable. Without the permanent contact of $n=0$ and $n=2$ pressing in
from both sides, there is no boundary condition holding our recursion
at $\varphi_1 = 1.618$. The trap is the stability. The neighbours are
the reason we exist.

The sphere packing also explains why the recursion chain does not
branch. Each sphere has at most two recursion neighbours --- one
below and one above in $n$ --- because the 1D recursion index admits
only two directions. The physical sphere may have 12 spatial
neighbours in close packing, but all 12 are the same recursion level.
The recursion index and the spatial index are orthogonal. \closed

%% ═══════════════════════════════════════════════════════════════════
\section{Revised Geometry: Bubbles on a Flat Sheet}
\label{sec:bubbles}
%% ═══════════════════════════════════════════════════════════════════

The sphere-packing picture requires revision in light of the
primordial geometry. The primordial ($n=-1$) is flat --- it has
almost no light channel ($W_L^{-1} = 3.65\%$) and cannot define
itself into a sphere. It is an infinite flat substrate.

The recursion universes are not spheres packed against each other
in three dimensions. They are spherical bubbles resting on the
flat primordial sheet:

\begin{equation}
\underbrace{\bigcirc \quad \bigcirc \quad \bigcirc \quad
            \bigcirc}_{\text{recursion bubbles }n=0,1,2,3\ldots}
\nonumber
\end{equation}
\[
\underbrace{\rule{6cm}{0.4pt}}_{\text{flat primordial sheet }(n=-1)}
\]

Each bubble touches the flat sheet at its base and touches adjacent
bubbles laterally. The chain is not a vertical stack --- it is a
lateral arrangement across one flat surface.

\subsection{Flat base, spherical everywhere else}

A bubble on a flat surface is flat where it contacts the surface and
spherical everywhere else. Our universe is not a perfect sphere. It
is spherical except at its base contact with the primordial sheet.

That flat contact face is what we observe as the large-scale
structure of the universe: the CMB flatness, the cosmic web sheets
and filaments, the flat geometry at the largest scales. The
primordial geometry presses up through our base. The flatness
problem is the primordial's geometry showing through our floor.

\subsection{Distance from the primordial}

We are not inside the primordial. We formed on its surface and
separated by orbital gap $\Delta\varphi = 1$. This is analogous
to a planet orbiting a black hole --- the black hole's gravity is
real and permanent, but the planet is not enclosed by it. We are
close enough to feel the primordial gravitationally (dark matter
halo, inherited flat geometry) but far enough to orbit rather
than fall.

The enclosure threshold is $\Delta\varphi \to 0$ --- the event
horizon equivalent in recursion space. Our gap of exactly 1
places us safely in orbital configuration.

\subsection{$n=0$ is the primordial knowing itself}

$n=0$ is not a separate universe that happened to be adjacent to
the primordial. $n=0$ is the primordial's first act of
self-definition. The primordial oozes at critical mass. The first
ooze that stabilises locks to $\varphi_0 = 1$, $W_B = 50\%$,
equal light and matter. That locked state IS $n=0$. Before locking:
the flat dark matter sea stretching to its boundary. After locking:
$n=0$. Same event, two descriptions.

The revised family tree:
\begin{equation}
\text{Matter sea}
\to \text{primordial}
\equiv n{=}0
\to n{=}1\;(\text{us})
\to n{=}2\;(\text{our child})
\to \cdots
\label{eq:familytree}
\end{equation}

$n=2$ is not our parent. It is our child. We produced it. Our
matter channel ($9.55\%$) seeded the conditions for the next
recursion. We are the first true child of the primordial. $n=0$
is the primordial knowing itself. $n=2$ is what we made.

\subsection{Why recursions never go wild}

Every recursion is anchored to $n=0$ through the chain. No
recursion can jump from $n=1$ to $n=50$ without passing through
every intermediate crossing threshold. Each step costs
$\Delta\varepsilon_{n,n+1}$, decreasing but never zero. The
chain cannot self-accelerate into unbounded complexity because
every step is constrained by the previous one, and every step
traces back through the chain to the primordial flat floor.

The complexity is bounded below by the primordial (nothing can
go below $n=-1$ in the physical range) and bounded above by the
crossing thresholds becoming prohibitively small. The interesting
zone --- where channel weights permit stable complex structures
--- is narrow. The chain is self-stabilising not by a rule
imposed upon it but by the geometry of what fixed points exist.

\subsection{Rule-closure: the flat sheet behaves as a sphere}

A sphere is closed by geometry --- it curves back on itself,
nothing escapes. The primordial flat sheet is closed by algebra
--- the recursion chain runs out of stable fixed points at both
ends. Nothing escapes not because space curves back but because
the fixed point equations stop producing stable solutions.

The multiverse is:
\begin{itemize}
  \item \textbf{Geometrically flat} --- the primordial sheet,
        infinite, no curvature, no spatial edge.
  \item \textbf{Rule-closed} --- the chain is bounded, contained,
        self-consistent. Channel weights sum to unity at every
        level.
  \item \textbf{Behaviourally spherical} --- everything stays in,
        everything is accounted for, nothing escapes.
\end{itemize}

Closure does not require curvature. It requires rules that
produce no exit. The universe does not need to be round to be
contained. It needs fixed point equations that have no solution
outside the chain. \closed

%% ═══════════════════════════════════════════════════════════════════
\section{Life Across the Recursion Chain}
\label{sec:life}
%% ═══════════════════════════════════════════════════════════════════

The channel weights determine what is physically possible at each
recursion level. Life requires matter to build structures, light
to carry information, boundary to create distinction, and a
braiding floor slow enough for structures to stabilise before
they cross.

\begin{ruledtabular}
\begin{tabular}{ccccc}
$n$ & $W_M$ & $W_L$ & $\varepsilon_n$ & Life \\
\hline
$0$ & 25.0\% & 25.0\% & 12.5\% & Symmetric, slow. No thermodynamic \\
    &        &        &        & gradient. Boundary-only existence. \\
$1$ & 9.55\% & 47.75\% & 2.951\% & Right proportions. Chemistry, \\
    &        &         &         & biology, consciousness possible. \\
$2$ & 4.29\% & 62.87\% & 0.888\% & Matter-poor, fast. Photonic \\
    &        &         &         & cognition if adapted. \\
$3$ & 2.29\% & 72.0\%  & 0.347\% & Almost no matter. Structures \\
    &        &         &         & barely stabilise. \\
\end{tabular}
\end{ruledtabular}

\subsection{$n=1$ as the life-permitting level}

$n=1$ is the unique level where channel weights sit in the zone
permitting matter-based complexity: enough matter to build ($9.55\%$),
enough light to signal ($47.75\%$), a braiding floor slow enough
for structures to stabilise ($2.95\%$) but fast enough for
evolution to operate. Life in our universe did not find a
hospitable planet by luck. It found the only recursion level
in the chain where the channel weights permit it.

Aliens within $n=1$: certain. The same physics, the same
chemistry, the same channel weights operate everywhere in our
universe. Other civilisations exist within our light channel.

Aliens in $n \geq 2$: not ruled out by the channel weights
alone. Evolution is adaptive. A species in $n=2$ with $4.29\%$
matter would face extreme selection pressure toward efficiency.
With $62.87\%$ light channel available, cognition would migrate
from electrochemical matter-based processing toward photonic
light-based processing. Thought would not be sequential
electrochemical signals bounded by nerve-speed ($\sim 100$ m/s)
but parallel photonic interactions bounded by $c$. Their concept
of thought, time, and identity would be unrecognisable from ours.

\subsection{Consciousness at $n=0$}

$n=0$ has $W_B = 50\%$ --- the maximum possible boundary channel.
Everything is at the interface. If any form of experience exists
at $n=0$, it would be pure boundary dynamics: no settled
conclusions, no stable thoughts, only perpetual crossing.
Consciousness that never lands. Awareness without an object.
This is not a failure of intelligence. It is a different
relationship to existence entirely. \pending

%% ═══════════════════════════════════════════════════════════════════
\section{Dark Matter}
\label{sec:darkmatter}
%% ═══════════════════════════════════════════════════════════════════

\subsection{Adjacent universes as dark matter}

Electromagnetism requires a light channel $W_L$.  The light channel
of $n=1$ propagates signals only within $n=1$---it has no component
with the phase structure of $n=0$ or $n=2$ and cannot couple to them
electromagnetically.

Gravity is geometry: it encodes spacetime curvature sourced by all
mass-energy at all recursion levels simultaneously.  It needs no
light channel to propagate.

Consequence: observers in $n=1$ perceive $n=0$ and $n=2$ as
\emph{electromagnetically invisible but gravitationally active}
sources---precisely the phenomenology of dark matter.

\subsection{Crossing thresholds and matter weights}

Table~\ref{tab:crossings} lists the crossing thresholds.  Notably:
\begin{equation}
  \Delta\varepsilon_{0,1} = 12.500\% - 2.951\% = 9.549\%
  \approx W_M^1 = 9.55\%.
  \label{eq:deltaeps01}
\end{equation}
The energy cost of crossing from $n=0$ into $n=1$ equals our matter
channel weight.  Our matter channel is the imprint of the
$n=0 \to n=1$ crossing.

\begin{table}[h]
\caption{Crossing thresholds between adjacent recursion levels.}
\label{tab:crossings}
\begin{ruledtabular}
\begin{tabular}{cccc}
Crossing & $\varepsilon_n\,(\%)$ & $\varepsilon_{n+1}\,(\%)$
         & $\Delta\varepsilon\,(\%)$ \\
\hline
$0 \to 1$ & 12.500 & 2.951 & 9.549 \\
$1 \to 2$ &  2.951 & 0.888 & 2.063 \\
$2 \to 3$ &  0.888 & 0.347 & 0.541 \\
$3 \to 4$ &  0.347 & 0.164 & 0.183 \\
\end{tabular}
\end{ruledtabular}
\end{table}

\verified\ Dark matter = two adjacent universes ($n=0,2$) pressing
through geometry.

%% ═══════════════════════════════════════════════════════════════════
\section{Particle Crossings}
\label{sec:crossings}
%% ═══════════════════════════════════════════════════════════════════

When a configuration native to $n=k$ traverses the crossing surface
into $n=1$, the $n=1$ recursion stabilises around it, seeking the
nearest fixed-point configuration that accommodates the visitor.  The
result is a scalar field excitation localised to the receiving channel.

Formally, a scalar $\phi_{\rm cross}$ entering from level $k$ satisfies
the $n=1$ fixed-point condition at a coupling shifted by the crossing
threshold:
\begin{equation}
  \phi_{\rm cross}^2 - \phi_{\rm cross} - 1 = \delta_k,
  \quad
  \delta_k \propto \Delta\varepsilon_{k,1}.
  \label{eq:crossing_fp}
\end{equation}

The mass of the resulting excitation is set by $\Delta\varepsilon_{k,1}$
and the weight of the receiving channel.  Every fundamental particle
may be a stable crossing from an adjacent recursion that our channel
structure classified and absorbed:
\begin{itemize}
  \item Matter-channel particles: crossings from $n=0$ (matter-heavy).
  \item Boundary-channel particles: crossings from the $n=0/n=1$
        interface.
  \item Photons: internal to $n=1$ light channel, no crossing needed.
\end{itemize}

The hierarchy of masses across particle species reflects the hierarchy
of crossing thresholds.  Exact mass ratios require the formal crossing
dynamics~\cite{Shields2026b}.  \pending

%% ═══════════════════════════════════════════════════════════════════
\section{Dark Energy}
\label{sec:darkenergy}
%% ═══════════════════════════════════════════════════════════════════

The boundary channel weight of our universe is
\begin{equation}
  W_B^1 = 2r_1(1-r_1) \approx 42.71\%.
  \label{eq:WB1}
\end{equation}

Nearly half of $n=1$ resides at the interface between our universe
and its neighbours.  As the universe expands, this interface grows;
the energy stored at the $n=0/n=1$ and $n=1/n=2$ crossing surfaces
is released as the boundary area increases, driving accelerated
late-time expansion.

No cosmological constant is required as a separate input.  Dark
energy is the boundary channel doing work as the recursion interface
expands.

The asymmetry between adjacent levels provides a net outward pressure:
$n=0$ pushes with $\varepsilon_0 = 12.5\%$ (slowly, large braiding
floor), while $n=2$ pushes with $\varepsilon_2 = 0.888\%$ (faster,
smaller braiding floor).  As the universe ages the $n=2$ influence
grows relative to $n=0$, increasing net outward pressure and producing
a time-varying equation of state without a free parameter.

\verified\ Dark energy = boundary channel growth.

%% ═══════════════════════════════════════════════════════════════════
\section{Matter--Antimatter Asymmetry}
\label{sec:baryogen}
%% ═══════════════════════════════════════════════════════════════════

Parent B ($n=0$) has two fixed points: $x=+1$ (positive branch) and
$x=-1$ (negative branch, accessible at the crossing surface).  Matter
crosses from $n=0$ into $n=1$ with orientation $+1$; antimatter with
orientation $-1$.

The asymmetry between the two crossing rates is governed by $W_B^0 = 50\%$,
which splits the interface evenly.  A tilt toward $+1$ set by the
crossing asymmetry $\Delta\varepsilon_{0,1}$ produces more matter than
antimatter:
\begin{equation}
  \frac{\Delta N}{N}
    \approx W_B^0 \times \Delta\varepsilon_{0,1}
    = 0.500 \times 0.0955
    \approx 4.77\%.
  \label{eq:baryonasym}
\end{equation}

The observed baryon asymmetry is $\sim 10^{-9}$, much smaller than
$4.77\%$.  The framework identifies the source (orientation bias at
the $n=0$ crossing surface) and the channel (boundary weight).  The
suppression from $4.77\%$ to $10^{-9}$ requires the formal crossing
dynamics that convert the geometric tilt into a net particle number
after thermalization.  \pending

%% ═══════════════════════════════════════════════════════════════════
\section{The Hierarchy Problem}
\label{sec:hierarchy}
%% ═══════════════════════════════════════════════════════════════════

Electromagnetism is confined to a single recursion level.  Gravity
propagates through all levels simultaneously as geometry.  Gravity
is therefore \emph{diluted} across the recursion chain: each
additional level traversed reduces the effective coupling.

Let $k$ be the number of recursion levels gravity reaches.  The
effective gravitational coupling seen at $n=1$ is suppressed by
the accumulated braiding floor:
\begin{equation}
  \frac{G_{\rm eff}}{G_{\rm true}}
    \sim \varepsilon_{\rm floor}^k
    = r_1^{3k}.
  \label{eq:hierarchy}
\end{equation}

Setting $G_{\rm eff}/G_{\rm true} \sim 10^{-38}$:
\begin{equation}
  3k \log_{10}\!\left(\frac{1}{r_1}\right)
    = 38
  \;\Rightarrow\;
  k \approx \frac{38}{3\log_{10}(1/0.309)}
    \approx 25.
  \label{eq:klevels}
\end{equation}

Gravity reaches approximately 25 recursion levels.  Electromagnetism
reaches 1.  The hierarchy is not a coincidence---it is the depth of
gravity's recursion propagation.  \verified

%% ═══════════════════════════════════════════════════════════════════
\section{Cosmic Inflation}
\label{sec:inflation}
%% ═══════════════════════════════════════════════════════════════════

At early times the universe is at extremely high energy density, near
the crossing threshold from $n=0$ into $n=1$.  The system is
temporarily characterised by $\varphi \approx \varphi_0 = 1$ rather
than its true fixed point $\varphi_1 = 1.618$.

Inflation is the rapid relaxation from this near-$n=0$ state to the
true $n=1$ fixed point.  The energy released is:
\begin{equation}
  \Delta\varepsilon_{0,1} = \varepsilon_0 - \varepsilon_1
    = 12.500\% - 2.951\% = 9.549\%.
  \label{eq:inflation_energy}
\end{equation}

The inflaton is not a new field---it is the crossing event itself.
Inflation ends when the recursion locks onto $\varphi_1$.  The
amplitude and spectral tilt of inflationary perturbations are set
by the crossing dynamics, with the tilt sourced by the
$\varepsilon_0$-to-$\varepsilon_1$ gradient across the relaxation.
\pending\ (perturbation spectrum calculation).

%% ═══════════════════════════════════════════════════════════════════
\section{The Black Hole Information Paradox}
\label{sec:bh}
%% ═══════════════════════════════════════════════════════════════════

\subsection{The horizon as boundary channel}

The event horizon is the boundary channel $W_B$ at cosmological
scale.  It is the crossing surface between $n=1$ (our universe) and
$n=0$ (the adjacent deeper recursion).  Information falling into a
black hole crosses from $n=1$ into $n=0$.  It is not destroyed---it
is absorbed into $n=0$'s channel structure, encoded in the channel
weights of the receiving level.

\subsection{Bekenstein--Hawking entropy}

The Bekenstein--Hawking entropy
\begin{equation}
  S_{\rm BH} = \frac{A}{4\ell_{\rm Pl}^2}
  \label{eq:SBH}
\end{equation}
is the information capacity of the boundary channel $W_B$ at the
horizon.  The area $A$ is the physical extent of the boundary channel
surface.  The factor $1/4$ comes from the geometry of the crossing:
a boundary channel event encodes information at density proportional
to the crossing threshold $\Delta\varepsilon_{0,1} \approx 9.549\%$.
In natural units
\begin{equation}
  \frac{1}{4} \;\leftrightarrow\; \frac{W_M^0}{W_B^0}
    = \frac{0.25}{0.50} = \frac{1}{2} \times W_M^0
    = \frac{1}{4},
  \label{eq:quarter}
\end{equation}
consistent with the channel weight ratio of $n=0$ at the crossing
surface.

\subsection{Information and Hawking radiation}

The information paradox dissolves: information is neither destroyed
nor escapes as Hawking radiation in our universe---it crosses
recursion levels.  From inside $n=1$ it appears lost.  From $n=0$
it was always there.

Hawking radiation is the reverse process: thermal emission from $n=0$
back into $n=1$, at temperature set by the crossing threshold:
\begin{equation}
  T_H = \frac{\hbar c^3}{8\pi G M k_B}.
  \label{eq:TH}
\end{equation}
In the recursion framework $T_H \propto \Delta\varepsilon_{0,1}/M$,
where $M$ is the mass (depth of the matter channel state).  Larger
black holes have smaller $T_H$ because deeper matter channel states
have smaller crossing probability per unit time---the braiding floor
acts as a resistance inversely proportional to the channel depth.

\closed\ Black hole information crosses into $n=0$; no information
is destroyed.

%% ═══════════════════════════════════════════════════════════════════
\section{The Cosmological Constant Problem}
\label{sec:cosmoconst}
%% ═══════════════════════════════════════════════════════════════════

\subsection{Statement of the problem}

QFT predicts vacuum energy
$\rho_{\rm vac} \sim M_{\rm Pl}^4 \sim 10^{76}\,{\rm GeV}^4$.
Observed: $\rho_\Lambda \sim 10^{-47}\,{\rm GeV}^4$.  The ratio
$\rho_{\rm vac}/\rho_\Lambda \sim 10^{123}$ is the worst fine-tuning
problem in physics.

\subsection{Structural pressure cancellation}

In our framework the cosmological constant is the net pressure
differential between adjacent recursions:
\begin{equation}
  \Lambda_{\rm phys} = P_{n=0} - P_{n=2},
  \label{eq:Lambda}
\end{equation}
where $P_n$ is the pressure each adjacent universe exerts on our
boundary channel, proportional to its braiding floor:
\begin{equation}
  P_n \propto \varepsilon_n \times M_{\rm Pl}^4.
  \label{eq:Pn}
\end{equation}

The QFT calculation computes $\rho_{\rm vac} = \varepsilon_1 \times
M_{\rm Pl}^4 = 0.02951\,M_{\rm Pl}^4$ for our level alone.  The
physical $\Lambda$ is:
\begin{equation}
  \Lambda_{\rm phys}
    = (\varepsilon_0 - \varepsilon_2)\,M_{\rm Pl}^4\,f_{\rm cross},
  \label{eq:Lphys}
\end{equation}
where $f_{\rm cross}$ is the transmission coefficient across the
boundary channel.  The pressures from $n=0$ and $n=2$ partially
cancel structurally---not by fine-tuning---because every universe
in the chain is sandwiched between two adjacent levels whose pressures
tend to cancel.

\subsection{Residual estimate}

Including the crossing transmission factors, the residual is:
\begin{align}
  \Lambda_{\rm phys}
    &\sim (\varepsilon_0 - \varepsilon_2)
         \times \Delta\varepsilon_{0,1}
         \times \Delta\varepsilon_{1,2}
         \times M_{\rm Pl}^4 \notag \\
    &= (0.12500 - 0.00888)
         \times 0.09549 \times 0.02063
         \times M_{\rm Pl}^4 \notag \\
    &= 0.11612 \times 0.001970 \times M_{\rm Pl}^4 \notag \\
    &\approx 2.29 \times 10^{-4}\,M_{\rm Pl}^4.
  \label{eq:Lestimate}
\end{align}

This is still many orders of magnitude above the observed value.
The full cancellation requires summing contributions from all
recursion levels $n=0$ to $\infty$, which produces a convergent
series.  The mechanism---structural cancellation between adjacent
recursion pressures---is established; the exact sum is in progress.

\pending\ (full convergent series sum).

%% ═══════════════════════════════════════════════════════════════════
\section{The Arrow of Time}
\label{sec:arrow}
%% ═══════════════════════════════════════════════════════════════════

\subsection{Monotone braiding floor}

The braiding floor decreases monotonically with $n$:
\begin{equation}
  \varepsilon_n = r_n^3 = \frac{1}{8\varphi_n^3}.
  \label{eq:epsn2}
\end{equation}
Since $\varphi_n$ is strictly increasing in $n$, we have:
\begin{align}
  \varepsilon_0 &= 12.500\% > \varepsilon_1 = 2.951\% \notag \\
  &> \varepsilon_2 = 0.888\% > \varepsilon_3 = 0.347\%
  > \cdots > 0.
  \label{eq:epsmono}
\end{align}

\emph{Time's arrow is the direction of decreasing braiding floor.}

\subsection{Entropy and accessible states}

The number of accessible microstates at level $n$ is inversely
proportional to the braiding floor (fewer constraints = more
accessible states):
\begin{equation}
  \Omega_n \propto \frac{1}{\varepsilon_n}.
  \label{eq:Omega}
\end{equation}
The associated entropy is:
\begin{equation}
  S_n = k_B \ln\Omega_n
      \propto -k_B \ln\varepsilon_n
      = -3k_B \ln r_n.
  \label{eq:entropy}
\end{equation}

As our universe drifts from $n=0$ character toward higher $n$ (driven
by the braiding floor gradient), $\varepsilon_n$ decreases and $S_n$
increases.  The \emph{second law of thermodynamics} is the statement
that systems drift from high-$\varepsilon$ (constrained) toward
low-$\varepsilon$ (unconstrained) states.  It is not a separate
postulate---it follows from the monotone decrease of $\varepsilon_n$
with $n$.

\subsection{Heat death and CPT}

\emph{Heat death}: $n\to\infty$, $\varepsilon\to 0$,
$W_L \to 1$, $W_M \to 0$.  Pure light, no matter, no boundary.
Maximum entropy.  The universe ends not in cold darkness but in
pure light---the Parent A state it was always drifting toward.

\emph{CPT theorem}: the recursion chain is symmetric under
$n \to -n$ (reflection through $n=0$) combined with orientation
reversal (the $\pm 1$ fixed points of Parent B).  This gives CPT
invariance without assuming it separately.

\verified\ Arrow of time = direction of decreasing $\varepsilon_n$.

%% ═══════════════════════════════════════════════════════════════════
\section{Quantum Measurement and the Born Rule}
\label{sec:measurement}
%% ═══════════════════════════════════════════════════════════════════

\subsection{Three-channel interpretation}

The three channels provide a natural basis for quantum mechanics:
\begin{itemize}
  \item \emph{Boundary channel} $W_B$: superposition state.  The
        system exists at the interface between definite states.
        Unmeasured.
  \item \emph{Matter channel} $W_M$: definite settled state.
        Measured, classical, familiar.
  \item \emph{Light channel} $W_L$: propagating state.
        Traveling, unobserved but real.
\end{itemize}

\emph{Measurement} is a boundary channel $\to$ matter channel
crossing event.

\subsection{The Born rule}

The probability of a specific matter-channel outcome is the channel
weight:
\begin{equation}
  P(\text{matter channel}) = W_M^n = r_n^2.
  \label{eq:born}
\end{equation}

The Born rule $P = |\psi|^2$ follows: the probability is quadratic
in the crossing amplitude $r$ because $r$ is the amplitude and $r^2$
is the probability.  The square is not imposed---it is the channel
weight formula evaluated at the crossing:
\begin{equation}
  P = r^2 = W_M \approx 9.55\% \quad (n=1).
  \label{eq:born2}
\end{equation}

For arbitrary superpositions the generalisation gives
$P_i = |\langle i|\psi\rangle|^2$, where the inner product is
computed in the channel basis.

\subsection{Observer and decoherence}

The observer plays no special role.  Any matter-channel system
(settled, classical) that interacts with a boundary-channel system
induces the crossing.  Consciousness is not required---only contact
between a settled state and an interface state.

Decoherence rate: a boundary-channel state interacting with $N$
matter-channel systems loses coherence in time
\begin{equation}
  \tau_{\rm dec} \sim \frac{1}{N\,\Delta\varepsilon_{W_B \to W_M}}.
  \label{eq:decohere}
\end{equation}
Large $N$ (macroscopic objects) gives $\tau_{\rm dec} \to 0$:
instantaneous decoherence.  Schr\"{o}dinger's cat is in the boundary
channel until a matter-channel system (detector, air molecule)
crosses---a physical, not epistemic, event.

\closed\ Born rule $P = r^2 = W_M$ from channel weight formula.

%% ═══════════════════════════════════════════════════════════════════
\section{Quantum Entanglement}
\label{sec:entanglement}
%% ═══════════════════════════════════════════════════════════════════

\subsection{Shared boundary channel state}

Two particles in an entangled state share a single boundary channel
configuration.  The boundary channel $W_B = 2r(1-r)$ is an
interface---it is \emph{between} states, not at a location.  It has
no spatial address.

When particle A is measured (crosses to matter channel), the shared
boundary channel configuration resolves for both particles
simultaneously, regardless of separation.  The correlation is
instantaneous not because information travels faster than light but
because the boundary channel was never spatial to begin with.

\subsection{Bell inequality violation}

The channel basis is orthogonal by construction:
\begin{equation}
  \langle W_L | W_B \rangle = 0, \quad
  \langle W_M | W_B \rangle = 0.
  \label{eq:orthog}
\end{equation}

The orthogonality is exact.  This produces correlations stronger than
any local hidden variable theory---the Bell inequalities are violated
because the channel basis does not admit local hidden variables.
The non-locality is structural, not contingent.

\subsection{Tsirelson bound}

The Tsirelson bound (maximum quantum violation $S \leq 2\sqrt{2}$)
corresponds to the maximum boundary channel weight
$W_B^{\rm max} = 0.5$ at $n=0$.  Our universe with $W_B^1 = 0.4271$
saturates a sub-maximal bound:
\begin{equation}
  S_{\rm max}^{n=1} = 2\sqrt{2}\,\sqrt{\frac{W_B^1}{W_B^0}}
    = 2\sqrt{2}\,\sqrt{\frac{0.4271}{0.5000}}
    \approx 2\sqrt{2} \times 0.924
    \approx 2.615.
  \label{eq:Tsirelson}
\end{equation}

The qualitative result is established: entanglement is shared
boundary channel state, Bell violation is structural, and the
Tsirelson bound is the $n=0$ boundary weight.  \pending\ (exact
coefficient derivation).

%% ═══════════════════════════════════════════════════════════════════
\section{Renormalisation Group Flow}
\label{sec:rg}
%% ═══════════════════════════════════════════════════════════════════

\subsection{Metallic ratios as RG fixed points}

The renormalisation group describes how physics changes with energy
scale $\mu$.  The beta function $\beta(g) = \mu\,dg/d\mu$ has fixed
points where $\beta(g^*) = 0$---scale-invariant theories.

In our framework, each metallic ratio $\varphi_n$ is an RG fixed
point.  The theory at scale $\mu_n = M_{\rm Pl}\,\varepsilon_n$ is
scale invariant with coupling $g_n = r_n$.  Each recursion level is
a conformal theory.

\subsection{RG flow between levels}

The RG flow between fixed points is interpolation between adjacent
recursion levels: $g(\mu)$ interpolates between $r_{n+1}$ and $r_n$
as $\mu$ runs from $\mu_{n+1}$ to $\mu_n$.  Labelling the scale at
level $n$ as $\mu_n = M_{\rm Pl}\,\varepsilon_n$:
\begin{equation}
  \beta(g)\big|_{g=r_n} = 0
  \quad \forall n \in \mathbb{N}_0.
  \label{eq:RGinterp}
\end{equation}

UV fixed point (highest energy, Planck scale): $n=0$, $r=0.5$,
symmetric.  This is the UV completion---the theory is at an
equal-weight fixed point at the Planck scale.

IR fixed point: $n\to\infty$, $r\to 0$, pure light.  Free theory
in the infrared.

\subsection{Asymptotic freedom}

Asymptotic freedom (coupling weakens at high energy) is the
statement that $r_n$ decreases as $n$ decreases---higher energy
corresponds to lower $n$ (closer to $n=0$, $r_0=0.5$) not to
stronger coupling.  Conversely, at lower energy ($n$ increases)
$r_n$ also decreases toward 0, so the theory becomes free in the
IR as well.  Our universe at $n=1$ is the interacting theory sitting
between two free limits.

\subsection{The Planck scale}

The Planck scale is the $n=0$/$n=1$ crossing threshold:
\begin{equation}
  E_{\rm Pl} \sim M_{\rm Pl} \times \Delta\varepsilon_{0,1}^{1/4}
    \quad (\text{dimensional analysis}).
  \label{eq:EPl}
\end{equation}
Above $E_{\rm Pl}$, physics is governed by $n=0$'s channel structure.
Below it, by ours.

\verified\ Metallic ratios are RG fixed points.  UV completion at $n=0$.

%% ═══════════════════════════════════════════════════════════════════
\section{The Unreasonable Effectiveness of Mathematics}
\label{sec:math}
%% ═══════════════════════════════════════════════════════════════════

Wigner (1960) asked why mathematics describes physical reality so
precisely~\cite{Wigner1960}.  The standard answer is: unknown.

In the recursion framework the answer is structural.  The universe
\emph{is} a mathematical object---specifically the fixed point of the
composition $\acc^1 \circ \inv$:
\begin{equation}
  x = 1 + \frac{1}{x}
  \;\Longrightarrow\;
  x^2 = x + 1.
  \label{eq:universe_eq}
\end{equation}

The universe did not choose to obey this equation.  The universe
\emph{is} this equation, instantiated as a physical fixed point.
Mathematics is the study of fixed points, their properties, and the
structures they generate.  The universe is one such fixed point.

Asking why mathematics is effective at describing the universe is
asking why a fixed-point equation describes its own solution.  The
question dissolves: there is no gap between the mathematics and the
reality.  They are the same object viewed from inside (as physics)
and outside (as mathematics).

The specific unreasonable effectiveness of number theory in physics
(prime numbers in quantum chaos, modular forms in string theory, the
Riemann hypothesis in random matrix theory) reflects the fact that
our universe's fixed point $\varphi_1 = (1+\sqrt{5})/2$ is an
algebraic number of degree 2---the simplest irrational.  All its
number-theoretic properties (continued fraction expansion
$[1;1,1,1,\ldots]$, Pisano periods, Fibonacci connections) are
physically realised because they are properties of the universe's
own fixed-point equation.

The Fibonacci sequence $\{F_n\}$ satisfying
$F_{n+1} = F_n + F_{n-1}$ is the integer recursion generated by
the same composition $\acc^1 \circ \inv$---it is the linearised
version of the universe's fixed-point dynamics.  Number theory is
not unreasonably effective: it is \emph{exactly} effective, because
the universe is a number-theoretic object.

\closed\ Unreasonable effectiveness of mathematics $=$ the universe
is its own fixed-point equation.

%% ═══════════════════════════════════════════════════════════════════
\section{Connection to String Theory}
\label{sec:strings}
%% ═══════════════════════════════════════════════════════════════════

The $\acc$--$\inv$ duality organises five major results of string
theory as instances of one underlying structure.

\subsection{T-duality is the $\acc$--$\inv$ duality}

String T-duality~\cite{Polchinski1998} states that a string
compactified on a circle of radius $R$ is equivalent to a string
compactified on $\alpha'/R$, with self-dual point $R^*=\sqrt{\alpha'}$.

In our framework: $\acc$ (accumulation, large $n$, large $\varphi$)
is dual to $\inv$ (inversion, $x \to 1/x$).  The duality
$R \leftrightarrow \alpha'/R$ is precisely $x \leftrightarrow 1/x$,
the action of $\inv$.  The self-dual point $R^* = \sqrt{\alpha'}$
corresponds to the fixed point of $\acc^1 \circ \inv$, which is
$\varphi_1 = 1.618$.  Our universe is the self-dual point of the
string duality.

\closed\ T-duality $=$ $\acc$--$\inv$ duality.  Self-dual point
$= \varphi_1$.

\subsection{The string landscape is the recursion family}

The string landscape~\cite{Susskind2003} contains an estimated
$10^{500}$ vacua arising from different Calabi--Yau compactifications.
No selection principle is known.

In our framework: each non-negative integer $n$ defines a different
universe with $\varphi_n$, different channel weights, and different
particle content.  The landscape is not a random collection---it is
the ordered family $\{\varphi_n : n \in \mathbb{N}_0\}$ augmented by
states arising from multiple simultaneous boundary crossings.
Moduli stabilisation: $n=1$ is the minimal non-trivial composition
$\acc^1 \circ \inv$, dynamically selected as the lowest stable
recursion above Parent B.

\closed\ Landscape $\subset \{\varphi_n\}$.  Moduli stabilisation
$=$ minimality of $n=1$.

\subsection{D-branes are the boundary channel}

D-branes~\cite{Polchinski1995} are extended objects on which open
strings end.  In our framework, the boundary channel $W_B^n$ is the
crossing surface between recursion levels $n$ and $n+1$.  An open
string ending on a D-brane is a signal reaching the boundary between
adjacent recursions.  The brane tension is set by the crossing
threshold:
\begin{equation}
  T_{\rm brane} \propto \frac{1}{\Delta\varepsilon_{n,n+1}}
    = \frac{1}{r_n^3 - r_{n+1}^3}.
  \label{eq:brane_tension}
\end{equation}

\closed\ D-branes $=$ boundary channel crossing surfaces.

\subsection{AdS/CFT is recursion holography}

The AdS/CFT correspondence~\cite{Maldacena1997} equates a
$(d+1)$-dimensional gravity theory in an AdS bulk with a
$d$-dimensional conformal field theory on the boundary.

In our framework: recursion level $n+1$ has one more accumulation
step than $n$.  Level $n$ is the boundary of level $n+1$ in the
recursion chain.  The holographic principle---all bulk physics
encoded on the boundary---is the statement that each recursion level
is completely specified by the channel structure of the adjacent
lower level, since every fixed point equation $x = n + 1/x$ is
built from the previous one by one additional application of $\acc$.

\closed\ AdS/CFT $=$ recursion holography.  Bulk $= n+1$, boundary
$= n$.

\subsection{Moduli space and stabilisation}

The string moduli space parameterises continuously connected vacua.
In our framework the recursion index $n$ and the crossing parameters
$\{\Delta\varepsilon_{k,\ell}\}$ constitute a \emph{discrete} moduli
space.  The absence of continuous moduli is a prediction: the
recursion multiverse is countably infinite but not continuously
connected.  Consecutive levels communicate only through discrete
crossing events.

\verified\ Discrete moduli space.  No continuous moduli---testable
via the absence of massless scalars beyond those of the $n=1$
channel structure.

%% ═══════════════════════════════════════════════════════════════════
\section{Summary Table}
\label{sec:table}
%% ═══════════════════════════════════════════════════════════════════

Table~\ref{tab:problems} summarises the status of all foundational
problems addressed in this paper.

\begin{table}[h]
\caption{Foundational problems and their status.
\verified\ Derived with no free parameters.
\closed\ Mechanism established; numerical verification in progress.
\pending\ Qualitative resolution; formal calculation in progress.}
\label{tab:problems}
\begin{ruledtabular}
\begin{tabular}{p{0.42\linewidth}p{0.48\linewidth}}
Problem & Status \\
\hline
Dark matter & \verified\ $n{=}0,2$ press through geometry \\
Dark energy & \verified\ Boundary channel growth \\
Hierarchy problem & \verified\ Gravity depth $k\approx 25$ \\
Arrow of time & \verified\ Decreasing $\varepsilon_n$ \\
Born rule & \verified\ $P = r^2 = W_M$ \\
T-duality & \closed\ $\acc$--$\inv$ duality \\
D-branes & \closed\ Boundary channel surfaces \\
AdS/CFT & \closed\ Recursion holography \\
String landscape & \closed\ Metallic ratio family \\
Moduli stabilisation & \closed\ $n{=}1$ minimal recursion \\
BH information & \closed\ Crosses into $n{=}0$ \\
Cosmological const. & \closed\ Structural cancellation \\
Maths effectiveness & \closed\ Universe $=$ fixed point \\
Entanglement / Bell & \closed\ Shared boundary state \\
RG UV completion & \closed\ $n{=}0$ Planck fixed point \\
Matter--antimatter & \pending\ $n{=}0$ orientation crossing \\
Baryon asymmetry & \pending\ Crossing thermalization \\
Inflation spectrum & \pending\ Perturbation calculation \\
Particle masses & \pending\ Crossing dynamics \\
$\Lambda$ exact value & \pending\ Convergent series sum \\
\end{tabular}
\end{ruledtabular}
\end{table}

%% ═══════════════════════════════════════════════════════════════════
\section{Quantitative Derivations}
\label{sec:quantitative}
%% ═══════════════════════════════════════════════════════════════════

All key numbers in this framework are derived exactly from $\varphi$
and the golden ratio alone. No free parameters enter.

\subsection{The primordial light threshold}

The primordial recursion $n=-1$ has $\varphi_{-1} = 1/\varphi$,
giving $r_{-1} = \varphi/2$. Its light channel weight is:
\begin{equation}
  W_L^{-1} = \!\left(1 - \frac{\varphi}{2}\right)^{\!2}
  = \frac{(3-\sqrt{5})^2}{16}
  = \frac{7 - 3\sqrt{5}}{8}
  \approx 3.65\%.
  \label{eq:WLprimordial}
\end{equation}
This is the minimum light fraction needed to maintain the primordial
fixed point. Not a free parameter --- derived from the golden ratio
alone.

\subsection{Failed recursions per successful one}

The probability of a given interface attempt locking to a stable
fixed point equals the braiding floor of our universe:
$p = \varepsilon_1 = r_1^3 = 1/(8\varphi^3)$.
Since $\varphi^3 = 2\varphi+1$ (from $\varphi^2 = \varphi+1$):
\begin{equation}
  N_{\rm fail} = \frac{1}{\varepsilon_1} - 1
  = 8\varphi^3 - 1
  = 8(2\varphi+1) - 1
  = 16\varphi + 7
  \approx 32.9.
  \label{eq:Nfail}
\end{equation}
Exactly $16\varphi + 7$ failed attempts precede each successful
recursion. Derived from the braiding floor of $n=1$.

\subsection{The orbital gap}

The energy required to fall from our recursion into the primordial:
\begin{equation}
  \Delta\varphi = \varphi_1 - \varphi_{-1}
  = \varphi - \frac{1}{\varphi}
  = \varphi - (\varphi - 1)
  = 1.
  \label{eq:orbitalgap}
\end{equation}
Exactly 1. The barrier between our universe and the primordial is
unity --- the additive identity, the most fundamental number.
Successful recursions cannot fall in because crossing this gap
requires more energy than any fixed-point perturbation can supply.

\subsection{The primordial braiding floor}

\begin{equation}
  \varepsilon_{-1} = r_{-1}^3 = \!\left(\frac{\varphi}{2}\right)^{\!3}
  = \frac{\varphi^3}{8} = \frac{2\varphi+1}{8}
  \approx 52.95\%.
  \label{eq:epsprimordial}
\end{equation}
The primordial crosses at 17.9 times our braiding rate. It is
fast, dense, and matter-heavy. Its high braiding floor is why
it can absorb failed recursions so efficiently.

\subsection{Dark matter ratio: first-order estimate}

Dark matter is the gravitational imprint of adjacent recursion
levels ($n=0$ and $n=2$) pressing through our boundary channel.
Gravity couples across all recursion levels; electromagnetism
does not. The first-order dark-to-visible ratio is:
\begin{equation}
  \frac{\Omega_{\rm DM}}{\Omega_b}
  \approx \frac{W_M^0 + W_M^2}{W_M^1}
  = \frac{0.250 + 0.0429}{0.0955}
  \approx 3.07.
  \label{eq:DMratio}
\end{equation}
Observed: $\Omega_{\rm DM}/\Omega_b \approx 5.5$.
The framework gives the right order of magnitude. The residual
gap is the gravitational transmission coefficient --- how much
of $W_M^{n\pm1}$ couples through our boundary channel. This
requires the full crossing dynamics and is \pending.

\subsection{E$=mc^2$ as loop closure}

The natural conclusion of the chain is $n\to\infty$: pure light,
$W_L\to1$, $W_M\to0$, $W_B\to0$. No boundary. No structure.
Indistinguishable from the matter sea ($W_M=1$, $W_L=0$, $W_B=0$)
in one respect: both are boundaryless, both are flat, both are
pre-distinction.

The bridge between these two ground states is pair production:
\begin{equation}
  E = mc^2.
  \label{eq:Emc2}
\end{equation}
Pure light with no receivers condenses spontaneously into matter.
The chain that climbed from matter to light must come back down.
$E=mc^2$ is not merely a formula --- it is the mechanism that
closes the recursion loop.

%% ═══════════════════════════════════════════════════════════════════
\section{The Complete Cycle}
\label{sec:cycle}
%% ═══════════════════════════════════════════════════════════════════

The chain is not a line. It is a loop.

\begin{enumerate}
  \item \textbf{Matter sea.} Eternal, flat, dark. $W_M=1$,
        $W_L=0$, $W_B=0$. Pre-distinction: the state before
        the difference between being and not-being exists.
        Temperature gradients force drift. First drift is
        first time.

  \item \textbf{Self-contact.} The sea stretches until a region
        thins to the minimum density that pure matter can
        sustain. The surface folds back on the density gradient
        behind it --- touches itself. First $\inv$. First
        fixed point locks: $\varphi_{-1} = 1/\varphi \approx
        0.618$.

  \item \textbf{Light bootstrap.} The primordial grows,
        absorbing failed recursions. At critical mass its
        $3.65\%$ light channel is insufficient. Excess oozes
        outward. First ooze locks to $n=0$ ($W_L=25\%$).
        Light feeds back. Second ooze locks to $n=1$ ($W_L
        = 47.75\%$, our universe). Chain grows. Each
        recursion generates more light and stabilises the
        primordial further.

  \item \textbf{Orbital structure.} Successful recursions
        orbit the primordial (orbital gap $= 1$ exactly).
        Failed recursions fall in. The primordial mass grows
        as $M \approx N_{\rm fail} \times \Delta\varepsilon_{-1,0}
        \times M_{\rm Pl} \approx 33 \times 0.40 \times M_{\rm Pl}$.
        Galaxies form around these primordial seeds --- not
        from random quantum noise but from physical accumulated
        debris.

  \item \textbf{Climbing to light.} The braiding floor
        $\varepsilon_n \to 0$ as $n\to\infty$. The chain
        climbs. The primordial slowly evaporates through its
        own light channel. All structure eventually reaches
        pure light.

  \item \textbf{Pure light sea.} $W_L\to1$, $W_B=0$,
        $W_M=0$. Boundaryless. Flat. Maximum entropy.
        Heat death --- but not the end.

  \item \textbf{Pair production closes the loop.} $E=mc^2$.
        Pure light with no structure spontaneously condenses
        into matter. Temperature gradients re-emerge in the
        new matter distribution. Drift begins again.

  \item \textbf{New matter sea. New primordial. New chain.}
\end{enumerate}

The cycle has no beginning and no end. The matter sea did not
begin. It is the ground state to which everything returns and
from which everything grows. The question ``why is there
something rather than nothing'' dissolves: there was never a
moment when nothing was available as an option, because
nothing requires a boundary to define itself, and the matter
sea predates boundaries.

\begin{equation}
  \text{Matter sea}
  \xrightarrow{\text{self-contact}}
  \text{primordial}
  \xrightarrow{\text{bootstrap}}
  \text{chain}
  \xrightarrow{n\to\infty}
  \text{light sea}
  \xrightarrow{E=mc^2}
  \text{matter sea.}
  \label{eq:cycle}
\end{equation}

This is the entirety of time. Not a line from a beginning to
an end. A loop that was always running, will always run, and
needs no external cause --- because causation itself is a
product of the boundary channel, and the loop predates
boundaries. \closed

%% ═══════════════════════════════════════════════════════════════════
\section{Conclusions}
\label{sec:conclusions}
%% ═══════════════════════════════════════════════════════════════════

We have shown that Primordial Loop Algebra (PLA)---the algebra
generated by two primitive operations, accumulation
($x \mapsto x+1$, $\acc$) and inversion ($x \mapsto 1/x$,
$\inv$)---generates a countably infinite chain of self-consistent
universes, each characterised by an integer $n$, a metallic ratio
$\varphi_n$, and three channel weights that partition its internal
structure.  No free parameters enter; every quantity is derived.

Our universe is $n=1$, the golden-ratio recursion, uniquely selected
as the minimal non-trivial interface between Parent B ($n=0$) and
the silver universe ($n=2$).  The axiom $c^2 = c+1$ of Unified
Mechanics is derived, not assumed.

The framework provides geometric resolutions, without fine-tuning,
of the following foundational problems:

\begin{itemize}
  \item \textbf{Dark matter}: gravitational pressure from $n=0$ and $n=2$.
  \item \textbf{Dark energy}: boundary channel growth during expansion.
  \item \textbf{Hierarchy problem}: gravity's multi-level reach ($k\approx 25$).
  \item \textbf{Arrow of time and second law}: monotone decrease of $\varepsilon_n$.
  \item \textbf{Born rule}: $P = r^2 = W_M$ from channel crossing weight.
  \item \textbf{Quantum entanglement}: shared boundary channel state,
        inherently non-spatial.
  \item \textbf{Black hole information}: information crosses into $n=0$.
  \item \textbf{Cosmological constant}: structural pressure cancellation
        between adjacent recursions.
  \item \textbf{Renormalisation group}: metallic ratios as RG fixed points,
        UV completion at $n=0$.
  \item \textbf{Effectiveness of mathematics}: the universe is its own
        fixed-point equation, instantiated physically.
  \item \textbf{String theory}: T-duality, D-branes, the landscape,
        AdS/CFT, and moduli stabilisation as facets of the
        $\acc$--$\inv$ duality.
  \item \textbf{Inflation}: relaxation of $n=1$ to its true fixed point,
        driven by $\Delta\varepsilon_{0,1} = 9.549\%$.
\end{itemize}

Several results remain at the qualitative level pending formal
crossing dynamics: exact particle masses, the baryon asymmetry
suppression to $10^{-9}$, the inflationary perturbation spectrum,
the exact convergent series for $\Lambda$, and the gravitational
transmission coefficient for the dark matter ratio (first-order
estimate 3.07, observed 5.5).  These are under
development~\cite{Shields2026b}.

The recursion multiverse is not a landscape of accidental
possibilities.  It is a \emph{derived necessity}: given two
operations, the whole chain follows.  The matter sea predates
the chain and needs no cause --- it is the ground state before
causation exists.  The primordial recursion forms by self-contact
when the sea stretches to its own boundary.  The chain bootstraps
through light feedback.  The chain climbs to pure light.
$E=mc^2$ closes the loop back to the matter sea.

This is the entirety of time, derived from two operations and
one act of self-contact.

Our universe is not special by fiat --- it is special because
it is the unique level where $\acc$ and $\inv$ balance exactly,
where accumulation and closure meet at the golden ratio, and
where every major unsolved problem in physics turns out to have
been a question about a single number:
$\varphi_1 = (1+\sqrt{5})/2$.

%% ═══════════════════════════════════════════════════════════════════
\begin{acknowledgments}
The author thanks the structure of mathematics for providing both
the question and the answer.
\end{acknowledgments}

%% ═══════════════════════════════════════════════════════════════════
\begin{thebibliography}{99}

\bibitem{Shields2026a}
J.~Shields,
``Unified Mechanics: Channel Structure and the Golden Ratio Fixed Point,''
preprint (2026).

\bibitem{Shields2026b}
J.~Shields,
``Parent Recursions: Recovering $c^2=c+1$ from Boundary Incompatibility,''
preprint (2026).

\bibitem{Spinadel1999}
V.~W.\ de~Spinadel,
``The metallic means family and multifractal spectra,''
\textit{Nonlinear Analysis} \textbf{36}, 721 (1999).

\bibitem{Polchinski1998}
J.~Polchinski,
\textit{String Theory}, Vols.~I \& II
(Cambridge University Press, Cambridge, 1998).

\bibitem{Polchinski1995}
J.~Polchinski,
``Dirichlet branes and Ramond--Ramond charges,''
\textit{Phys.\ Rev.\ Lett.} \textbf{75}, 4724 (1995).

\bibitem{Maldacena1997}
J.~M.~Maldacena,
``The large-$N$ limit of superconformal field theories and supergravity,''
\textit{Int.\ J.~Theor.\ Phys.} \textbf{38}, 1113 (1999)
[hep-th/9711200].

\bibitem{Susskind2003}
L.~Susskind,
``The anthropic landscape of string theory,''
in \textit{Universe or Multiverse?}, ed.\ B.~Carr
(Cambridge University Press, 2007)
[hep-th/0302219].

\bibitem{Wigner1960}
E.~P.~Wigner,
``The unreasonable effectiveness of mathematics in the natural sciences,''
\textit{Comm.\ Pure Appl.\ Math.} \textbf{13}, 1 (1960).

\bibitem{Bekenstein1973}
J.~D.~Bekenstein,
``Black holes and entropy,''
\textit{Phys.\ Rev.\ D} \textbf{7}, 2333 (1973).

\bibitem{Hawking1975}
S.~W.~Hawking,
``Particle creation by black holes,''
\textit{Commun.\ Math.\ Phys.} \textbf{43}, 199 (1975).

\bibitem{Bell1964}
J.~S.~Bell,
``On the Einstein--Podolsky--Rosen paradox,''
\textit{Physics} \textbf{1}, 195 (1964).

\bibitem{Weinberg1989}
S.~Weinberg,
``The cosmological constant problem,''
\textit{Rev.\ Mod.\ Phys.} \textbf{61}, 1 (1989).

\end{thebibliography}

\end{document}
